### 2020-06-11 paper.tex
The DPM paper and literature for CoronaApp

\documentclass{llncs}

\usepackage{fancyhdr}
\usepackage{flushend}
\usepackage{amsfonts,amssymb,amsmath,alltt}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{color}

\usepackage{makeidx}
\pagestyle{plain}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage[english]{babel}

\renewcommand{\ttdefault}{cmtt}
%This version DOES NOT suppress line breaks
% newcommands etc
\newenvironment{ttbox}{\begin{alltt}\ttbraces\small\tt}%
                      {\end{alltt}}
%the new definition of \. suppresses line breaks
\def\ttbraces{\let\.=\nobreak\chardef\{=`\{\chardef\}=`\}\chardef\|=`\\}

\newcommand{\symb}[1]{\makebox{\it #1}} 
\newcommand\ie{i.e.\!\,, }
\newcommand\rel{Re${\cal{L}}$}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand\ttand{\mbox{{$\land$}}}
\newcommand\ttor{\mbox{{$\lor$}}}
\newcommand\ttcap{\mbox{{$\cap$}}}
\newcommand\ttcup{\mbox{{$\cup$}}}
\newcommand\ttfun{\mbox{{$\Rightarrow$}}}
\newcommand\ttmimp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttimp{\mbox{{$\longrightarrow$}}}
\newcommand\ttequiv{\mbox{{$\equiv$}}}
\newcommand\ttexists{\mbox{{$\exists$}}}
\newcommand\ttforall{\mbox{{$\forall$}}}
\newcommand\ttneg{\mbox{{$\neg$}}}
\newcommand\ttneq{\mbox{{$\neq$}}}
\newcommand\ttin{\mbox{{$\in$}}}
\newcommand\ttnin{\mbox{{$\notin$}}}
\newcommand\ttImp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttlam{\mbox{\( \lambda \)}}
\newcommand\tttimes{\mbox{\( \times \)}}
\newcommand\ttlbrack{\mbox{\(\llbracket\)}}
\newcommand\ttrbrack{\mbox{\( \rrbracket \)}}
\newcommand\noie{\textit{i.e.},\xspace}
\newcommand\eg{~\textit{e.g.},\xspace}
\newcommand\noeg{\textit{e.g.},\xspace}
\newcommand\ttatI{\mbox{\( @_G \)}}
\newcommand\ttto[1]{\mbox{{$\to^{#1}$}}}
\newcommand\ttleq{\mbox{{$\le$}}}
\newcommand\ttrelI{\mbox{{$\to_{i}$}}}
\newcommand\ttalpha{\mbox{{$\alpha$}}}
\newcommand\tttau{\mbox{{$\tau$}}}
\newcommand\ttsubseteq{\mbox{{$\subseteq$}}}
\newcommand\ttf{\mbox{{$f$}}}
\newcommand\ttvdash{\mbox{{$\vdash$}}}
\newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
\newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
\newcommand\ttattand[1]{\mbox{{$\oplus_{\wedge}^{#1}$}}}
\newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
\newcommand\ttrelIstar{\mbox{{$\to_{i}^*$}}}

\begin{document}
\frontmatter

\mainmatter
\title{Modeling Corona App in Isabelle}
\author{Florian Kamm\"uller and Bianca Lutz}

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bialut@gmail.com}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the currently designed
Corona App.
\end{abstract}

\section{Introduction}
\label{sec:intro}

\bibliographystyle{abbrv}
\bibliography{../insider}

\end{document}

### 2020-07-01 paper.tex
Intro started and structure laid.

@@
 \mainmatter
-\title{Modeling Corona App in Isabelle}
+\title{Modeling and analysing the Coronavirus warning app with the Isabelle Infrastructure framework}
 \author{Florian Kamm\"uller and Bianca Lutz}
@@
 \begin{abstract}
-We provide a model in the Isabelle Infrastructure framework of the currently designed
-Corona App.
+We provide a model in the Isabelle Infrastructure framework of the recently published
+Coronavirus warning app. The app supports breaking infection chains by informing users
+whether they have been in close contact to an infected person. The app has a decentralized
+architecture that supports anonimity of users.
+We provide a formal model of the existing app with the Isabelle Infrastructure framework
+to show up some natural attacks in a very abstract model. We then use the security
+refinement process of the Isabelle Infrastructure frmework to highlight how the use of
+continuosuly changing ephemeral ids improves the anonymity.
 \end{abstract}
@@
 \label{sec:intro}
+The German Chancelor Angela Merkel has strongly supported the publication of
+the mobile phone Corona warning app by publicly proclaiming that the ``Corona
+App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
+in Germany have downloaded the app with 6 million on the first day.
+This app is one amongst many similar application that aim at the very important goal
+to ``break infection chains'' by providing timely information of users whether they
+have been exposed to close contact with a person that has been infected.
+
+The Coronavirus warning app has taken a long time to develop being published only on
+16th June 2020. It was a quite costly project but this was mainly due to the management of
+Telekom and SAP being in the driving seat. But the app has been designed with great
+attention on privacy:  a distributed architecture \cite{} has been adopted after a long and
+heated debate with supporters of a central architecture. The distributed architecture is
+based on a very clever distributed application design whereby users phones are sending
+highly anonymised so called ``Ephemeral IDs'' at physical locations via the Bluetooth
+protocol. The app saves those IDs of people in close proximity. When at a later date an
+infected person reports his infection to a central server, the unique root ID is published
+and in the daily check all mobile phones connecting to the central server can download
+the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
+all Ephemeral IDs that have been saved over the last 14 days allow users phones to
+regularly check whether their user has been exposed to an infected person and issue
+a warning to the user. The warning issued by the Corona warning app entitles to
+ahving a Corona test done (which at the time of writing is not normally possible).
+
+The Isabelle Infrastructure framework \cite{kam:20a} allows modeling and analysing
+architecture and scenarios including physical and logical entities, actors, and policies
+within the interactive theorem prover Isabelle supported with temporal locic, Kriple
+structures, and attack trees. It has been applied for example to Insider analysis in
+airplanes \cite{kam:20b}, privacy in IoT healthcare \cite{kam:18b}, and recently also
+to blockchain protocols \cite{kn:20}.
+
+\TODO{Motivation: Why bother re-engineering a formal specification for a nicely developed privacy oriented app? physical aspects (adoption rate), formal proof etc}
+
+In this paper, we first provide some background in Section \ref{sec:background}:
+we give a detailed overview of the development history and security and privacy
+relevant parts of the Coronavirus warning app (Section \ref{sec:history})
+and some essential facts about the Isabelle Infrastructure framework
+(Section \ref{sec:isainf}).
+We then present our model (Section \ref{sec:model}) and analysis of privacy and
+attacks (Section \ref{sec:ana}) before drawing some conclusions (Section \ref{sec:concl}).
+
+The formal model in the Isabelle insider framework is fully mechanised and proved in
+Isabelle (sources available \cite{kam:18smc}).
+
+\section{Background}
+\label{sec:background}
+
+\subsection{History of Decentralized Archicture}
+\label{sec:history}
+
+\subsection{Isabelle Infrastructure framework}
+\label{sec:isainf}
+
+\section{Model}
+\label{sec:model}
+
+\section{Analysis}
+\label{sec:ana}
+
+\section{Conclusions}
+\label{sec:concl}
 
 \bibliographystyle{abbrv}
-\bibliography{../insider}
+\bibliography{insider}
 

### 2020-07-02 paper.tex
Fixed the import of ModTrans. Now, refinement statement is ok.

@@
 \mainmatter
-\title{Modeling and analysing the Coronavirus warning app with the Isabelle Infrastructure framework}
+\title{Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework}
 \author{Florian Kamm\"uller and Bianca Lutz}
@@
 We provide a model in the Isabelle Infrastructure framework of the recently published
-Coronavirus warning app. The app supports breaking infection chains by informing users
+Corona-virus warning app. The app supports breaking infection chains by informing users
 whether they have been in close contact to an infected person. The app has a decentralized
-architecture that supports anonimity of users.
+architecture that supports anonymity of users.
 We provide a formal model of the existing app with the Isabelle Infrastructure framework
 to show up some natural attacks in a very abstract model. We then use the security
-refinement process of the Isabelle Infrastructure frmework to highlight how the use of
-continuosuly changing ephemeral ids improves the anonymity.
+refinement process of the Isabelle Infrastructure framework to highlight how the use of
+continuously changing ephemeral ids improves the anonymity.
 \end{abstract}
@@
 \label{sec:intro}
-The German Chancelor Angela Merkel has strongly supported the publication of
+The German Chancellor Angela Merkel has strongly supported the publication of
 the mobile phone Corona warning app by publicly proclaiming that the ``Corona
@@
 
-The Coronavirus warning app has taken a long time to develop being published only on
+The Corona-virus warning app has taken a long time to develop being published only on
 16th June 2020. It was a quite costly project but this was mainly due to the management of
@@
 based on a very clever distributed application design whereby users phones are sending
-highly anonymised so called ``Ephemeral IDs'' at physical locations via the Bluetooth
+highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
 protocol. The app saves those IDs of people in close proximity. When at a later date an
@@
 a warning to the user. The warning issued by the Corona warning app entitles to
-ahving a Corona test done (which at the time of writing is not normally possible).
+having a Corona test done (which at the time of writing is not normally possible).
 
-The Isabelle Infrastructure framework \cite{kam:20a} allows modeling and analysing
+The Isabelle Infrastructure framework \cite{kam:20a} allows modeling and analyzing
 architecture and scenarios including physical and logical entities, actors, and policies
-within the interactive theorem prover Isabelle supported with temporal locic, Kriple
+within the interactive theorem prover Isabelle supported with temporal logic, Kripke
 structures, and attack trees. It has been applied for example to Insider analysis in
@@
 we give a detailed overview of the development history and security and privacy
-relevant parts of the Coronavirus warning app (Section \ref{sec:history})
+relevant parts of the Corona-virus warning app (Section \ref{sec:history})
 and some essential facts about the Isabelle Infrastructure framework
@@
 
-The formal model in the Isabelle insider framework is fully mechanised and proved in
+The formal model in the Isabelle insider framework is fully mechanized and proved in
 Isabelle (sources available \cite{kam:18smc}).
@@
 
-\subsection{History of Decentralized Archicture}
+\subsection{History of Decentralized Architecture}
 \label{sec:history}

### 2020-07-02 paper.tex
added a shoert seciton on Motivcation to Intro

@@
 
-\TODO{Motivation: Why bother re-engineering a formal specification for a nicely developed privacy oriented app? physical aspects (adoption rate), formal proof etc}
+%{Motivation: Why bother re-engineering a formal specification for a nicely developed privacy oriented app? physical aspects (adoption rate), formal proof etc}
+The technical advantage of modeling an application in the Isabelle Insider framework lies
+in (a) having explicit representations of infrastructures, actors and policies in a
+formal model that (b) allows additional automated verification of security properties
+within the interactive theorem prover Isabelle.
+Although the Corona-virus app hass been produced based on a sophisticated security
+concept conceived by experts in the field, to our knowledge, no formal verification
+has been involved.
 

### 2020-07-02 paper.tex
fadded to motivationsection in Intro

@@
 within the interactive theorem prover Isabelle.
-Although the Corona-virus app hass been produced based on a sophisticated security
+Although the Corona-virus app has been produced based on a sophisticated security
 concept conceived by experts in the field, to our knowledge, no formal verification
-has been involved.
+has been involved. Even if a ``post-production'' formal specification seems pointless,
+it allows to reveal weak points of the architecture, show that the measures that have
+been conceived are suitable to cover those weak points. Thereby, we believe
+that our current work is useful to increase the trust in the Corona-virus app necessary for
+its wide adoption which in turn is crucial for it to be efficient.
 

### 2020-07-03 paper.tex
inserted some adapted short version of Isabelle Infrastructure desciption. Redrew the diagram in draw.io

@@
 \label{sec:isainf}
+\TODO{Adapt: this is copied from FMBC paper}
+The Isabelle Infrastructure is built in the interactive generic theorem prover
+Isabelle/HOL \cite{npw:02}. As a framework, it supports formalisation and proof of
+systems with actors and policies. It originally emerged from verification of insider 
+threat scenarios but it soon became clear that the theoretical concepts, like temporal
+logic combined with Kripke structures and a generic notion of state transitions were
+very suitable to be combined with attack trees into a formal security engineering process
+\cite{suc:16} and framework \cite{kam:19a}.
+
+Figure \ref{fig:theorystruc} gives an overview of the Isabelle Infrastructure 
+framework with its layers of object-logics -- each level below embeds the one
+above showing the novel contribution of this paper in blue on the top. 
+\begin{figure}
+\begin{center}
+  \includegraphics[scale=.5]{theory_structure.pdf}
+\end{center}
+%\vspace{-.5cm}
+\caption{Generic Isabelle Infrastructure framework applied to Corona-warning app.}
+\label{fig:theorystruc}
+\end{figure}
+The formal model of the Corona-warning app uses the Isabelle Infrastructure framework instantiating it
+by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
+to physical locations. The Ephermeral IDs, their sending and change is added to Infrastructures
+by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
+for the actions move, get, and put. The details of the newly adapted Infrastructure are
+presented in Section \ref{sec:model}.
+Technically, an Isabelle theory file \texttt{Infrastructure.thy} builds on top of the theories for Kripke 
+structures and CTL (\texttt{MC.thy}), attack trees (\texttt{AT.thy}), and security refinement 
+(\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
+for IBC, express relevant and interesting properties and conduct interactive proofs (with the 
+full support of the powerful and highly automated proof support of Isabelle).
+The IBC theory itself is an adaptation of the Infrastructure theory of the 
+Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
+In the remainder of this paper, we introduce the model that we conceived for IBC. 
+All Isabelle sources are available online \cite{kam:20gitsc}.
 

### 2020-07-03 paper.tex
added a refinement section copied from arxive paper and needs to be  merged in. Also adapted the intro and background on Isabelle Infrastructures.

@@
 \newcommand\ttrelI{\mbox{{$\to_{i}$}}}
+\newcommand\ttrelIstar{\mbox{{$\to^*$}}}
+\newcommand\ttrel[1]{\mbox{{$\to_{#1}$}}}
+\newcommand\ttrelstar[1]{\mbox{{$\to_{#1}^*$}}}
 \newcommand\ttalpha{\mbox{{$\alpha$}}}
+\newcommand\ttsupseteq{\mbox{{$\supseteq$}}}
+\newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
 \newcommand\tttau{\mbox{{$\tau$}}}
@@
 \newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
-\newcommand\ttrelIstar{\mbox{{$\to_{i}^*$}}}
+\newcommand\ttsigma{\mbox{{$\sigma$}}}
+\newcommand\ttmref[1]{\mbox{{$\sqsubseteq_{#1}$}}}
+\newcommand\ttmeref{\ttmref{\mathcal{E}}}
+\newcommand\ttsn{\mbox{{\texttt{s}$_0$}}}
+\newcommand\ttsigmap{\mbox{{$\sigma'$}}}
+\newcommand\ttecal{\mbox{$\mathcal{E}$}}
+\newcommand\ttimg{\mbox{$\triangleleft$}}
 
@@
 
+The contributions of this paper are (a) formal re-engineering of the Corona-virus warning app
+(b) providing an additional security and privacy analysis with interactice theorem proving
+certification of a novel view on the system architecture including actors, locations, and policies,
+(c) an formal definition of a security refinement process that allows to improve a system
+based on attacks found by the attack tree analysis.
+
 In this paper, we first provide some background in Section \ref{sec:background}:
@@
 We then present our model (Section \ref{sec:model}) and analysis of privacy and
-attacks (Section \ref{sec:ana}) before drawing some conclusions (Section \ref{sec:concl}).
+attacks (Section \ref{sec:ana}). The found attck on the first abstract specification
+motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
+framework is introduced and illustrated on the Corona-virus warning app (Section \ref{sec:ref})
+before drawing some conclusions (Section \ref{sec:concl}).
 
@@
 \label{sec:isainf}
-\TODO{Adapt: this is copied from FMBC paper}
+%\TODO{Adapt: this is copied from FMBC paper}
 The Isabelle Infrastructure is built in the interactive generic theorem prover
@@
 
+\subsubsection{Overview}
+\label{sec:isainfover}
 Figure \ref{fig:theorystruc} gives an overview of the Isabelle Infrastructure 
@@
 \end{figure}
+
+\subsubsection{Instantiation of Framework}
 The formal model of the Corona-warning app uses the Isabelle Infrastructure framework instantiating it
@@
 (\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
-for IBC, express relevant and interesting properties and conduct interactive proofs (with the 
-full support of the powerful and highly automated proof support of Isabelle).
-The IBC theory itself is an adaptation of the Infrastructure theory of the 
-Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
-In the remainder of this paper, we introduce the model that we conceived for IBC. 
+for the Corona-virus warning app, express relevant and interesting properties, and conduct
+interactive proofs (with the full support of the powerful and highly automated proof support
+of Isabelle).
+%The Corona-virus warning app theory itself is an adaptation of the Infrastructure theory of the 
+%Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
+%In the remainder of this paper, we introduce the model that we conceived for . 
 All Isabelle sources are available online \cite{kam:20gitsc}.
+
+\subsubsection{Refinement}
+An interesting feature that has been integrated into the Isabelle Infrastructure
+framework motivated by security engineering formal specifications for IoT healthcare
+system is  is an extension of the formal specifciation process introducing
+refinement of Kripke structures \cite{kam:19a,kam:20a}. It refines a system model based on a 
+formal definition of a combination of trace refinement and structural 
+refinement (or datatype refinement). The definition allows to prove property preservation results 
+crucial for an iterative development process.
+The refinements of the system specification  can be interleaved with attack 
+analysis while security properties can be proved in Isabelle. In each iteration
+security qualities are accumulated while continuously attack trees scrutinize
+the design.
 
@@
 \label{sec:ana}
+
+\section{Refinement}
+\label{sec:ref}
+\TODO{This section is just copied over from arxive paper - adapt}
+Intuitively, a refinement changes some aspect of the type of
+the state, for example, replaces a data type by a richer datatype or
+restricts the behaviour of the actors. The former is expressed directly 
+by a mapping of datatypes, the latter is incorporated into the state
+transition relation of the Kripke structure that corresponds to the 
+transformed model.
+%This relationship between the Kripke structure can be better understood
+%visually (see Figure \ref{fig:modtrans})
+In other words, we can encode a refinement within our framework
+as a relation on Kripke structures that is parametrized additionally by
+a function that maps the refined type to the abstract type.
+The direction ``from refined to abstract'' of this type mapping may seem 
+curiously counter-intuitive. However, the actual refinement is given by the 
+refinement that uses this function as an input. The refinement
+then refines an abstract to a more concrete system specification. 
+The additional layer for the refinement 
+%(the red box in Figure \ref{fig:theorystruc})
+can be formalised in Isabelle as a
+%second\footnote{The first refinement 
+%relation in this framework is on attack trees summarized in Section \ref{sec:at}.} 
+refinement relation 
+$\sqsubseteq_{\mathcal{E}}$. 
+The relation \texttt{mod\_trans} is typed as a relation over triples --
+a function from a threefold Cartesian product to \texttt{bool}, the 
+type containing true and false only.  
+The type variables $\sigma$ and $\sigma'$ input to the type constructor 
+\texttt{Kripke} represent the abstract state type and the concrete state type. 
+Consequently, the middle element of the triples selected by the relation 
+\texttt{mod\_trans} is a function of type $\sigma' \Rightarrow \sigma$ 
+mapping elements of the refined state to the abstract state.
+The expression in quotation marks after the type is again the
+infix syntax in Isabelle that allows the definition of mathematical notation
+instead of writing \texttt{mod\_trans} in prefix manner. This nicer infix
+syntax is already used in the actual definition.
+Finally, the arrow \texttt{\ttImp} is the implication of Isabelle's  meta-logic
+while $\ttimp$ is the one of the {\it object} logic HOL. 
+They are logically equivalent but of different
+types: within a HOL formula $P$, for example, as below $\forall x. P \ttimp Q$, only the implication $\ttimp$
+can be used.
+\begin{ttbox}
+ mod_trans ::  (\ttsigma Kripke \tttimes (\ttsigma' \ttfun \ttsigma) \tttimes \ttsigma' Kripke) 
+               \ttfun bool                  ("_ \ttmref{(\_)} _")
+  K \ttmeref K' \ttequiv \ttforall s' \ttin states K'. \ttforall s \ttin init K'. 
+             s \ttrelstar{\sigma'} s' \ttimp \ttecal(s) \ttin init K 
+             \ttand \ttecal(s) \ttrelstar{\sigma} \ttecal(s')
+\end{ttbox}
+The definition of \texttt{K \ttmeref\, K'} states that for any state $s$ 
+of the refined Kripke structure that can be reached by the state transition
+in zero or more steps from an initial state $s_0$ of the refined Kripke 
+structure, the mapping ${\mathcal E}$ from the refined to the abstract 
+model's state must preserve this reachability, i.e., the image of
+$s_0$ must also be an initial state and from there the image of $s$
+under ${\mathcal E}$ must be reached with $0$ or $n$ steps.
+
+\subsection{Property Preserving System Refinement}
+A first direct consequence of this definition is the following lemma
+where the operator \texttt{$\ttimg$} in \texttt{\ttecal\ttimg(init K')}
+represents function image, that is the set, $\{\ttecal(x). x \in \texttt{init K'}\} $.
+\begin{ttbox}
+{\bf{lemma}} init_ref: K \ttmeref K' \ttImp \ttecal\ttimg(init K') \ttsubseteq init K
+\end{ttbox}
+A more prominent consequence of the definition of refinement 
+is that of property preservation. Here, we show that refinement preserves the
+CTL property of ${\sf EF} s$ which means that a reachability property true in the
+refined  model \texttt{K'} %has also been 
+is already true in the abstract model.
+A state set $s'$ represents a property %since we use 
+in the predicate transformer view of properties as sets of states. 
+The additional condition on initial states ensures that we cannot ``forget'' them. 
+%initial states.
+%The theorem states that, if the 
+%property \texttt{{\sf EF} s'} is true in the concrete model its  has been true 
+%can be reached from the initial state in 
+%\texttt{K'} then it is also possible  to reach the image of this property
+% in the abstract Kripke structure \texttt{K}.
+\begin{ttbox}
+{\bf{theorem}} prop_pres: 
+   K \ttmeref K'  \ttImp init K \ttsubseteq \ttecal\ttimg(init K') \ttImp
+   \ttforall s' \ttin Pow(states K'). K' \ttvdash {\sf EF} s' 
+              \ttimp K \ttvdash {\sf EF} (\ttecal\ttimg(s'))
+\end{ttbox}
+It is remarkable, that our definition of refinement by Kripke 
+structure refinement entails property preservation and makes it possible 
+to prove this as a theorem in Isabelle once for all, i.e., as a meta-theorem.
+However, this is due to the fact that our generic definition of state transition
+allows to explicitly formalise such sophisticated concepts like reachability.
+For practical purposes, however, the proof obligation of showing that
+a specific refinement is in fact a refinement is rather complex
+justly because of the explicit use of the transitive closure of the state
+transition relation.
+In most cases, the refinement will be simpler. Therefore, we offer
+additional help by the following theorem that uses a stronger characterisation
+of Kripke structure refinement and shows that our refinement follows
+from this.
+\begin{ttbox}
+{\bf{theorem}} strong_mt: 
+\ttecal\ttimg(init K') \ttsubseteq init K \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') 
+\ttImp K \ttmeref K'
+\end{ttbox}
+This simpler characterisation is in fact a stronger one: we could have $s \ttrel{\sigma'} s'$ 
+in the refined Kripke structure \texttt{K'} and $\neg(\ttecal(s) \ttrel{\sigma} \ttecal(s'))$
+but neither $s$ nor $s'$ are reachable from initial states in \texttt{K'}.
+For cases, where we want to have the simpler one-step proviso but still need 
+reachability we provide a slightly weaker version of \texttt{strong\_mt}.
+\begin{ttbox}
+{\bf{theorem}} strong_mt':  
+\ttecal\ttimg(init K') \ttsubseteq init K \ttand (\ttexists s0 \ttin init K'. s0  \ttrelIstar s)
+ \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') \ttImp K \ttmeref K'
+\end{ttbox}
+
+This idea of property preservation coincides with the classical idea of
+trace refinement as it is given in process algebras like CSP. In this view,
+the properties of a system are given by the set of its traces. Now, a refinement
+of the system is given by another system that has a subset of the traces of the 
+former one.
+Although the principal idea is similar, we greatly extend it since our notion
+additionally incorporates refinement. Since we include a state map 
+\texttt{\ttsigma'\ttfun \ttsigma} in our refinement map, we additionally
+allow structural refinement: the state map generalises the basic idea of
+trace refinement by traces corresponding to each other but allows additionally
+an exchange of data types. 
+As we see in the application to the case study, the refinement steps may
+sometimes just specialise the traces: in this case the state map 
+\texttt{\ttsigma'\ttfun \ttsigma} is just identity.
 

### 2020-07-07 notes-paper-dpm2020.tex
Bits and pieces: Related works, protocol description, motivation, attacks ...

\documentclass{article}
\usepackage{xcolor}
\usepackage{ulem}

\definecolor{styC}{rgb}{1,0,1}    %magenta
\definecolor{nsrC}{HTML}{0F8282}  %teal
\definecolor{noteC}{rgb}{0,0,1}   %blue
\definecolor{todoC}{HTML}{FE5010} %orange

\newcommand*{\TODO}[2][todoC]{{\color{#1} #2}}
\newcommand*{\TODOsty}[2][styC]{\TODO[#1]{#2}}
\newcommand*{\TODOnsr}[2][nsrC]{\TODO[#1]{#2}}
\newcommand*{\NOTE}[2][noteC]{\TODO[#1]{\textit{#2}}}
\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}

\usepackage{fancyhdr}
\usepackage{flushend}
\usepackage{amsfonts,amssymb,amsmath,alltt}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{graphicx}

\usepackage{makeidx}
\pagestyle{plain}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage[english]{babel}

\begin{document}

%%%%%
\section{Legend}
Dear Florian,

here's how to read this.\medskip\\
%%

I use the following color code to indicate different levels of incompleteness or uncertainty (usually with regard to the language, not the facts):
\begin{itemize}
\item \TODOnsr{\underline{Not sure}}\\
  Doesn't sound weird exactly but some doubt lingers and I'm simply not sure
  (punctuation, grammar, wording ...).
\item \TODOsty{\underline{Pretty sure it's wrong or could be better, at least}}\\
  Might be poor style or I don't know which word to pick or something else
  that feels not quite right/appropriate.
\item \NOTE{\underline{Comments}}\\
  Notes, remarks and such like
\item \TODO{\underline{TODO}}\\
  Something to check or fix. Mostly bibtex references (see footnotes starting with ``\TODO{\textbf{BIB-REF}:}'').
\end{itemize}

If there's anything to say beyond the color, I put it in a footnote. These footnotes are colored the same way the corresponding phrase is and bracketed to set them apart from ordinary footnotes -- there's indeed at least one of those hidden somewhere ;D...

Section headings try to summarize what a passage is (mainly) concerned with -- think of them as kind of working titles.
The sections are in no particular order.\medskip\\

Hope you can get anything out of it -- in parts it's rather notes than ready-to-use texts (and less than I expected/hoped for) :/

Use it, call me or send me an email in case you have any questions, send me a picture of scribbles you put on a print out and I'll send you (hopefully) ready-to-use revisions of stuff you want to use, whatever suits you ... just let me know :)\medskip\\

LG\bigskip\\
PS: I honestly don't think the development of the German corona app took remarkably long -- I seem to remember reading something like that in your introduction.
Judging from my daily (professional) experience I'm sort of
amazed they got it done so \textit{fast}, i.e. almost in time ;) -- sadly enough, I really mean that...
I don't doubt it was costly, though.

%%%%%
\section{Related works: DP-3T and PEPP-PT}
%
This paper \TODOnsr{is mainly concerned} with the \TODOnsr{protocol architecture} proposed by the \textit{Decentralized Privacy-Preserving Proximity Tracing} project (DP-3T).
The main reason to focus \TODOnsr{on} this particular family of protocols is the \textit{Exposure Notification Framework}, \TODOsty{jointly published} by Apple and Google, \TODOsty{following/that follows} a number of core principles of the DP-3T proposal. This API is not only utilized \TODOsty{in/by} the German \textit{Corona-Warn-App} (CWA) but has the potential of being widely adopted in future app developments \TODOsty{(we might see/that might emerge)}, due to the reach of \TODOsty{players like Apple and Google}.

There are, however, competing architectures noteworthy, namely protocols developed under the roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT), e.g. PEPP-PT-ROBERT\TODOref{ROBERT github repository} and PEPP-PT-NTK\TODOref{NTK ???; mentioned in ``Response to Analysis of DP-3T''}.

Neither DP-3T nor PEPP-PT \TODOnsr{is synonym for} just a single protocol. Each project endorses different protocols with unique properties in terms of privacy and data protection.\\
Yet, on \TODOsty{a more abstract level/a higher level of abstraction}, it seems \TODOnsr{feasible} to distinguish two basic architectures: Protocols \TODOnsr{as} endorsed by PEPP-PT might be characterized as centralized architectures whereas DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
  \TODOsty{As we will see}, DP-3T involves a central backend server. It is decentralized with regard to \TODOsty{collection and evaluation} of contact information:
  In centralized architectures the server provides a risk scoring services, whereas decentralized approaches rely on local risk assessment and, thus, do not need to share contact information with the \TODOnsr{backend}.}

\NOTE{%
  \vspace{.25em}\\
  ``DP-3T is a free-standing effort, originally started at EPFL and ETHZ.''\\
  Some of its members participated in the PEPP-PT project but have since
  resigned from the initiative.\TODOfn[noteC]{%
    github:DP-3T/documents\#april-8th-2020-the-relationship-between-dp-3t-and-pepp-pt
  }}

%%%%%
\section{Basic DP-3T protocol: How does it work? How does it relate to ENF/CWA? (broad strokes)}
%
Upon installation\TODOnsr{,} the app generates secret daily seeds to derive so called \textit{Ephemeral IDs} (EphIDs) \TODOnsr{from}. EphIDs are generated locally with cryptographic methods and cannot be connected to one another but only reconstructed \TODOnsr{from} the \TODOsty{secret seed they were derived from/corresponding secret seed}.\\
During normal operation each client broadcasts \TODOnsr{their} EphIDs via Bluetooth \TODOsty{whilst}\TODOfn[styC]{Too pompous? Is ``while scanning...'' better?} scanning \TODOnsr{for} EphIDs broadcasted by other devices in the vicinity. Collected EphIDs are stored locally along with associated metadata such as signal attenuation and date. In DP-3T the contact information gathered is never shared but only evaluated \TODOsty{locally/on the device}.\\
After patients are diagnosed \TODOnsr{(officially)}\TODOnsr{,} they are entitled to upload specific data to a central backend server. This data is accumulated by the \TODOnsr{backend}\TODOfn[nsrC]{Is ``backend'' a valid ellipsis or is it ``backend server''?} and redistributed to all clients regularly to \TODOnsr{provide the means for} local risk \TODOsty{scoring/assessment}, i.\,e. determining whether collected EphIDs match \TODOnsr{those} broadcasted by \TODOnsr{now-confirmed} COVID-19 patients during the last \TODOsty{e.\,g. 14}\TODOfn[styC]{%
      I can't figure out how to express that it is some fixed number of days that might be 14 but doesn't have to be (without a lot of lengthy blahblah or getting all formal using some $N$ and ``where ...'').} %
days.

In the most simple (and insecure) protocol proposed by DP-3T \TODOref{Low-cost decentralized proximity tracing; DP-3T Whitepaper p. 14ff} this \TODOsty{basically} translates into publishing the daily seeds used to derive EphIDs \TODOnsr{from}.
\TODOsty{Aside from additional security measures}\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with this level of detail). But this sounds just sh...} like signing content the server provides \TODOnsr{as a general rule}, the protocol implemented by ENF and, hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve \TODOsty{privacy and data protection properties} to different degrees but are more \TODOsty{costly/complicated}\TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.

\NOTE{%
  \vspace{.25em}\\
  I have a nice drawing to illustrate this (including some possible mitigation techniques outlined by CWA and DP-3T, respectively).
  I'll scan it when I'm in the office this week.}

%%%%%
\section{Formal verification using HOL: Why bother?}
%
\NOTE{%
  Sorry, loose notes is all I got on this one...}
%

\begin{itemize}
\item Quote\TODOref{DP-3T Whitepaper p. 2}:
  ``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]''\smallskip\\
  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the lines of:}
  Despite strong claims with regard to mathematical support, there is, as of yet, no formal verification (known to the authors). \TODOsty{So we just thought, we throw our hat in the ring.}
\item Explore different concepts of refinement: Maybe it is possible to pin down (i.e. exemplify) different concepts of refinement (data refinement, action refinement, trace refinement (aka spec refinement?) ...) -- and perhaps combinations thereof -- with concrete attack scenarios. I'm sure there'd be something to learn from that (esp. with respect to IsabelleAT).\\
  \NOTE{Sorry, can't specify it any further :(}
\end{itemize}

%%%%%
\section{What about attacks? Which of them do we consider? \sout{Why is that?}}
%
There is a variety of \TODOsty{interesting/noteworthy} privacy and security issues that \TODOsty{justify/deserve} \TODOsty{formal consideration} (and verification). The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of \TODOsty{interesting material} in terms of attack scenarios and possible mitigation strategies, e.g. \TODOref{Analysis of DP-3T ($\rightarrow$ github:PEPP-PT); Response ($\rightarrow$ github:DP-3T); PSRE and (perhaps) Analysis of PEPP-PT by DP-3T (somewhere on github:DP-3T, I guess; links can be found in Response to Analysis of DP-3T)}.

\NOTE{\\%
  ``Analysis of DP-3T'' for example addresses the following attack scenarios
  (categories of attacks):
  \begin{itemize}
  \item False Alert Injection Attacks
    \begin{itemize}
    \item Backend Impersonation / False Report
    \item Replay and Relay attacks
    \end{itemize}
  \item \TODOsty{Deanonymization/Tracking}\TODOfn[styC]{%
    Not sure you could really call it ``Deanonymization attack''. I'm not even sure ``deanonymization'' is a proper word...}
    Attacks (the authors don't call it that)
    \begin{itemize}
    \item Based on establishing a connection between collected EphIDs and
      (target) users, an attacker might:
      \begin{itemize}
      \item Track user's positions by tracking their Bluetooth signal
        (as recognized by the EphIDs they broadcast).\\
        $\rightarrow$ \textit{Location Tracking}
      \item Disclose a user's (positive) test status.\\
        $\rightarrow$ \textit{Test Status Disclosure}
      \item Disclose private encounters.
      \end{itemize}
    \end{itemize}
  \end{itemize}
}

\TODOsty{We focus on what we call Deanonymization Attacks.}\medskip\\

Although addressed in ``Analysis of DP-3T'', the two attacks we're concerned with, are classified by DP-3T as generic and inherent risks, respectively\TODOref{Privacy and Security Risk Evaluation of Digital Proximity Tracing Systems: GR 5 and IR 1}, \sout{and, hence, only \TODO{briefly} \TODOsty{addressed}} \TODO{$\leftarrow$ no they're not:} \NOTE{PSRE actually contains a quite thorough analysis of how infected individuals might be identified. (With pretty much the same conclusion I drew.)\\
  \label{PSRE}
  I only realized yesterday, what a treasure trove PSRE really is
  (or would have been)...
  Most (if not all) of the stuff I was thinking about the last couple of weeks
  is noted in there... Apparently I didn't read the right document :/ ...
  To be honest, this is a bit discouraging...\\
  See: Privacy and Security Risk Evaluation of Digital Proximity Tracing
  Systems, p. 5f -- IR 1: Identifying infected individuals}

%%%
\subsection{Why choose Bluetooth Beacons as starting point for building a formal representation?}
\begin{itemize}
\item Simple and thus easy to formalize.
\item Deanynomization attacks start here.
\item The Bluetooth Low Energy Beacon mechanism is employed in DP-3T and PEPP-PT protocols \TODOnsr{alike}.\\
  \NOTE{Perhaps the only thing they can agree on ;)}\smallskip\\
  There is one significant difference, though:
  \begin{itemize}
  \item DP-3T:
    The architecture requires that reported users can be recognized
    by the apps of their encounters
    (to some extend; e.\,g. mitigation by cuckoo filters).\smallskip\\
    Quote \TODOref{Analysis of DP-3T, p. 10}:
    ``Based on that, the only way to mitigate this attack is to deny the same
    privileges as the apps to individuals.''\smallskip\\
    \NOTE{(IMHO) peculiar English; even so, a direct quote}
  \item PEPP-PT:
    The backend server issues EphIDs and never shares any secret seeds,
    i.\,e. deanynomizing EphIDs \TODOnsr{arguably} requires more effort
    (= breaking into the server).
    \TODOnsr{In return}, more sensitive data (= contact time data)
    is shared with the server\TODOnsr{, to begin with}.
  \end{itemize}
\end{itemize}

%%%
\subsection{Deanonymization Attacks}
%%
\subsubsection{Different goals, different attacks, different levels of mitigation}
Regularly changing EphIDs might prevent location tracking effectively, or, \TODOnsr{at least, limit it to an acceptable degree}. While \TODOsty{(possibly)} having \TODOnsr{little to no effect}\TODOfn[nsrC]{Maybe that's too much; better (?): ``significantly/much less impact/effect on''} \TODOsty{on/in} scenarios where \TODOsty{the goal is to learn if} some particular user was infected.

\NOTE{%
  \vspace{.25em}\\
  \underline{Idea:}\\
  In order for location tracking to succeed, a valid/active EphID (i.\,e. one
  that is actually/currently broadcasted) is required.
  As a consequence, the time frame in which (additional) location information
  can be gathered is limited by the frequency at which EphIDs change.\\
  A positive test result, on the other hand, can be reported a (fixed)
  maximum number of days after the compromised EphID was used
  and still yield the desired information.
  This maximum time frame is independent of any frequency at which EphIDs
  may change.}

\NOTE{%
  \vspace{.25em}\\
  I have two (pretty high level, similar -- i.\,e. comparable) drawings
  depicting Location Tracking and Test Status Disclosure, respectively.\\
  Could help to illustrate this
  ``strategy X works in this context pretty good,
  but in that one not so much'' idea.}

%%
\subsubsection{Tracking someone's location}
\TODOnsr{If} an attacker can connect EphIDs to specific users or devices, he can track their whereabouts: For as long as a deanonymized EphID is valid, i.\,e. is broadcasted by the respective client, scanning for this id would suffice to locate \TODOnsr{this} client.\\
The range of Bluetooth reception can be improved by \TODOsty{sufficiently good/ high-performance} antennas. While \TODOnsr{stationary set up} hardware \TODOsty{can free an attacker of the need for actual, physical presence}.

%%
\subsubsection{Revealing someone's (positive) test status}
Like any notification about \TODOsty{e.\,g. a STD}\TODOfn[styC]{%
  Again, I don't know if you can use ``e.\,g.'' like that}
  might inform a spouse about the infidelity of the other, given the right circumstances, being notified about a possible COVID-19 exposure might reveal illicit information to the recipient like who it was \TODOsty{that} tested \TODOnsr{positive}.\\
  This is an inherent vulnerability of any such \TODOnsr{(notification)} system\TODOnsr{,
    whether it is analogous or} \TODOsty{digital/digitally aided}.
  The likelihood of said circumstances to \TODOnsr{occur}, however, might increase when notifications where carried out automatically.\\
  This is a \TODOnsr{quite interesting question to investigate}; its formal treatment could benefit considerably from the flexibility IsabelleAT offers: The ability to incorporate aspects of the physical world is \TODOsty{surely of special interest in this context}\TODOfn[styC]{Lame and awkward!}.

  \NOTE{\smallskip\\%
    DP-3T:PSRE\TODOref[noteC]{%
      Privacy and Security Risk Evaluation of Digital Proximity Tracing Systems:
      IR 1, p. 5f}
    contains a detailed analysis of how infected individuals might be identified.
    In the end, however, this attack --
    being (rightfully) classified as inherent risk -- is dismissed:
    ``These weaknesses are both minor and a generic problem in all
    designs based on Bluetooth proximity detection.''\TODOref[noteC]{%
      Response to Analysis of DP-3T:
      I: s.5.2 Deanonymizing Known Reported User, p. 2}\medskip\\
    If there was a chance of quantifying quality of support
    (i.\,e. a meaningful, yet abstract concept of effectiveness
    of proximity tracing in terms of aiding traditional contact tracing)...
    The question of ``cost'' (e.\,g. in the sense of increased risk)
    could be interesting, indeed.\\
    Too bad, this quantification business seems rather unlikely.}
  \pagebreak\\
  Quote \TODOref{%
    Privacy and Security Risk Evaluation of Digital Proximity Tracing Systems:
    IR 1, p. 6}:
  ``We emphasize that these attacks work against any contact tracing system,
  as they rely on the core proximity tracing functionality:
  notifying at-risk people.
  Without this notification, proximity tracing system would be useless.
  The risk is inherent to proximity tracing.''\medskip\\
  %
  Suppose, however, proximity tracing could be restricted to places
  of particular interest, e.\,g. places where ``mass gatherings'' are
  more likely to occur:
  public transport, supermarkets and such.\\
  \TODOsty{Maybe (probably?) this would be worse, privacy-wise,
    maybe it would be beneficial.}
  In either way, such a system wouldn't be useless.\\
  How much usefulness really stands to loose
  (by any digital proximity tracing system) is a question
  that still needs answering anyway.
  \NOTE{(As far as I know -- but that's not very far, to be honest)}
\end{document}

### 2020-07-07 paper.tex
integrated Chapter 2 and 3 of Nainca's notes

@@
 \newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
+\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
+\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
 \newcommand\ttand{\mbox{{$\land$}}}
@@
 
-The Corona-virus warning app has taken a long time to develop being published only on
-16th June 2020. It was a quite costly project but this was mainly due to the management of
+The Corona-virus warning app
+%has taken a long time to develop being published only on
+%16th June 2020. It
+was a quite costly project but this was mainly due to the management of
 Telekom and SAP being in the driving seat. But the app has been designed with great
@@
 In this paper, we first provide some background in Section \ref{sec:background}:
-we give a detailed overview of the development history and security and privacy
-relevant parts of the Corona-virus warning app (Section \ref{sec:history})
-and some essential facts about the Isabelle Infrastructure framework
-(Section \ref{sec:isainf}).
-We then present our model (Section \ref{sec:model}) and analysis of privacy and
+we give a detailed overview of the development history %and security and privacy
+ajnd relevant parts of the Corona-virus warning app (Section \ref{sec:history})
+We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
+Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
 attacks (Section \ref{sec:ana}). The found attck on the first abstract specification
@@
 
-\section{Background}
-\label{sec:background}
+\section{Background and related work}
 
-\subsection{History of Decentralized Architecture}
+\subsection{DP-3T and PEPP-PT}
 \label{sec:history}
+This paper is mainly concerned with the architecture and protocols proposed by the
+Decentralized Privacy-Preserving Proximity Tracing project (DP-3T).
+The main reason to focus on this particular family of protocols is the
+\textit{Exposure Notification Framework}, jointly published by Apple and Google, that adopts
+the main core principles of the DP-3T proposal. This API is not only used in the German
+Corona-virus warning app but has the potential of being widely adopted in future
+app developments that might emerge due to the reach of players like Apple and Google.
+
+There are, however, competing architectures noteworthy, namely protocols developed under the
+roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT), e.g.
+PEPP-PT-ROBERT \TODOref{ROBERT github repository} and PEPP-PT-NTK\TODOref{NTK ???; mentioned in
+  ``Response to Analysis of DP-3T''}.
+
+Neither DP-3T nor PEPP-PT are synonymous for just one single protocol. Each project endorses
+different protocols with unique properties in terms of privacy and data protection.
+Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
+protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
+DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
+  As we will see, DP-3T involves a central backend server. It is decentralized with regard to
+  collection and evaluation of contact information:
+  In centralized architectures the server provides a risk scoring services, whereas decentralized
+  approaches rely on local risk assessment and, thus, do not need to share contact information with
+  the backend.}
+
+\subsubsection{Basic DP-3T protocol: How does it work? How does it relate to ENF/CWA? (broad strokes)}
+%
+Upon installation,the app generates secret daily seeds to derive so called \textit{Ephemeral IDs}
+(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
+to one another but only reconstructed from the secret seed they were derived from.
+
+During normal operation each client broadcasts its EphIDs via Bluetooth whilst scanning for
+EphIDs broadcasted by other devices in the vicinity. Collected EphIDs are stored locally along
+with associated metadata such as signal attenuation and date. In DP-3T the contact information
+gathered is never shared but only evaluated locally (on the device).
+
+If patients are tested positive on the Corona-virus, they are entitled to upload specific data to a
+central backend server. This data is accumulated by the backend server and redistributed to all
+clients regularly to provide the means for local risk scoring, i.\,e., determining whether collected
+EphIDs match those broadcasted by now-confirmed Corona-virus patients during the last, e.\,g., 14, 
+days.
+
+In the most simple (and insecure) protocol proposed by DP-3T
+\TODOref{Low-cost decentralized proximity tracing; DP-3T Whitepaper p. 14ff}
+this basically translates into publishing the daily seeds used to derive EphIDs from.
+Aside from additional security measures,
+%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
+like signing content the server provides as a general rule, the protocol implemented by ENF and,
+hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
+privacy and data protection properties to different degrees but are more costly.
+%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
 

### 2020-07-07 paper.tex
udapte

@@
 regularly check whether their user has been exposed to an infected person and issue
-a warning to the user. The warning issued by the Corona warning app entitles to
-having a Corona test done (which at the time of writing is not normally possible).
+a warning to the user. The user then needs to contact health autohorities.
+%The warning issued by the Corona warning app entitles to
+%having a Corona test done (which at the time of writing is not normally possible).
 
@@
 \section{Background and related work}
-
+\label{sec:background}
 \subsection{DP-3T and PEPP-PT}
@@
 
-\subsubsection{Basic DP-3T protocol: How does it work? How does it relate to ENF/CWA? (broad strokes)}
+\subsubsection{Basic DP-3T protocol: How does it work?}
+%  How does it relate to ENF/CWA? (broad strokes)}
 %
-Upon installation,the app generates secret daily seeds to derive so called \textit{Ephemeral IDs}
+Upon installation, the app generates secret daily seeds to derive so called \textit{Ephemeral IDs}
 (EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
@@
 \label{sec:concl}
+\subsection{Protection goals of attacks}
+
+\subsection{Relevance of the approach}
 

### 2020-07-08 paper.tex
added some more text from Bianca

@@
 \label{sec:ref}
-\TODO{This section is just copied over from arxive paper - adapt}
+%\TODO{This section is just copied over from arxive paper - adapt}
 Intuitively, a refinement changes some aspect of the type of
@@
 
+In addition, as we can observe in the following example in the second refinement
+step, we also have a simple implicit version of {\it action refinement}. In an
+action refinement, traces may be refined by combining consecutive system events
+into atomic events thereby reducing traces. This will be detailed on the example
+below.
+
+\section{Refining the Specification}
+\label{sec:corref}
+
 \section{Conclusions}

### 2020-07-08 paper.tex
small changes

@@
 infected person reports his infection to a central server, the unique root ID is published
-and in the daily check all mobile phones connecting to the central server can download
+and in the daily check all mobile phones connecting to the central server download
 the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
-all Ephemeral IDs that have been saved over the last 14 days allow users phones to
+all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
 regularly check whether their user has been exposed to an infected person and issue
-a warning to the user. The user then needs to contact health autohorities.
+a warning to the user and recommendation to contact health authorities.
 %The warning issued by the Corona warning app entitles to
@@
 formal model that (b) allows additional automated verification of security properties
-within the interactive theorem prover Isabelle.
-Although the Corona-virus app has been produced based on a sophisticated security
-concept conceived by experts in the field, to our knowledge, no formal verification
-has been involved. Even if a ``post-production'' formal specification seems pointless,
+with CTL, Kripke structures and Attack Trees within the interactive theorem prover
+Isabelle.
+
+The Corona-virus app has been produced based on a sophisticated security
+concept conceived by experts in the field, %to our knowledge, no formal verification
+%
+and has strong claims with regard to mathematical support:
+
+``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
+%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
+However, there has been, as of yet, to our knowledge no formal verification been involved.
+Even if a ``post-production'' formal specification seems pointless, 
 it allows to reveal weak points of the architecture, show that the measures that have
@@
 The contributions of this paper are (a) formal re-engineering of the Corona-virus warning app
-(b) providing an additional security and privacy analysis with interactice theorem proving
+(b) providing an additional security and privacy analysis with interactive theorem proving
 certification of a novel view on the system architecture including actors, locations, and policies,
-(c) an formal definition of a security refinement process that allows to improve a system
-based on attacks found by the attack tree analysis.
+(c) a formal definition of a security refinement process that allows to improve a system
+based on attacks found by the attack tree analysis and (d) an application of the refinement to
+improve security of the Corona-virus app specification.
 
@@
 we give a detailed overview of the development history %and security and privacy
-ajnd relevant parts of the Corona-virus warning app (Section \ref{sec:history})
+and relevant parts of the Corona-virus warning app (Section \ref{sec:history})
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
 Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
-attacks (Section \ref{sec:ana}). The found attck on the first abstract specification
+attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
 motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
@@
 EphIDs broadcasted by other devices in the vicinity. Collected EphIDs are stored locally along
-with associated metadata such as signal attenuation and date. In DP-3T the contact information
+with associated meta-data such as signal attenuation and date. In DP-3T the contact information
 gathered is never shared but only evaluated locally (on the device).
@@
 The Isabelle Infrastructure is built in the interactive generic theorem prover
-Isabelle/HOL \cite{npw:02}. As a framework, it supports formalisation and proof of
+Isabelle/HOL \cite{npw:02}. As a framework, it supports formalization and proof of
 systems with actors and policies. It originally emerged from verification of insider 
@@
 by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
-to physical locations. The Ephermeral IDs, their sending and change is added to Infrastructures
+to physical locations. The Ephemeral IDs, their sending and change is added to Infrastructures
 by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
@@
 \subsubsection{Refinement}
-An interesting feature that has been integrated into the Isabelle Infrastructure
+An additional feature that has been integrated into the Isabelle Infrastructure
 framework motivated by security engineering formal specifications for IoT healthcare
-system is  is an extension of the formal specifciation process introducing
+system is an extension of the formal specification process introducing
 refinement of Kripke structures \cite{kam:19a,kam:20a}. It refines a system model based on a 
@@
 In other words, we can encode a refinement within our framework
-as a relation on Kripke structures that is parametrized additionally by
+as a relation on Kripke structures that is parameterized additionally by
 a function that maps the refined type to the abstract type.
@@
 %(the red box in Figure \ref{fig:theorystruc})
-can be formalised in Isabelle as a
+can be formalized in Isabelle as a
 %second\footnote{The first refinement 
@@
 However, this is due to the fact that our generic definition of state transition
-allows to explicitly formalise such sophisticated concepts like reachability.
+allows to explicitly formalize such sophisticated concepts like reachability.
 For practical purposes, however, the proof obligation of showing that
@@
 In most cases, the refinement will be simpler. Therefore, we offer
-additional help by the following theorem that uses a stronger characterisation
+additional help by the following theorem that uses a stronger characterization
 of Kripke structure refinement and shows that our refinement follows
@@
 \end{ttbox}
-This simpler characterisation is in fact a stronger one: we could have $s \ttrel{\sigma'} s'$ 
+This simpler characterization is in fact a stronger one: we could have $s \ttrel{\sigma'} s'$ 
 in the refined Kripke structure \texttt{K'} and $\neg(\ttecal(s) \ttrel{\sigma} \ttecal(s'))$
@@
 \texttt{\ttsigma'\ttfun \ttsigma} in our refinement map, we additionally
-allow structural refinement: the state map generalises the basic idea of
+allow structural refinement: the state map generalizes the basic idea of
 trace refinement by traces corresponding to each other but allows additionally
@@
 As we see in the application to the case study, the refinement steps may
-sometimes just specialise the traces: in this case the state map 
+sometimes just specialize the traces: in this case the state map 
 \texttt{\ttsigma'\ttfun \ttsigma} is just identity.

### 2020-07-08 paper.tex
added one more comment of Bianca on potential of theoretical study of refinement

@@
 heated debate with supporters of a central architecture. The distributed architecture is
-based on a very clever distributed application design whereby users phones are sending
+based on a very clever distributed application design whereby users' phones are sending
 highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
@@
 the design.
+One of the contributions of this paper is to explore different concepts of refinement:
+the formal expression of refinemnet, enables to pin down (i.e. exemplify) different concepts
+of refinement (data refinement, action refinement, trace refinement (aka spec refinement) and
+combinations thereof with concrete attack scenarios.
 

### 2020-07-08 paper.tex
added the conclusion subjection 7.2

@@
 One of the contributions of this paper is to explore different concepts of refinement:
-the formal expression of refinemnet, enables to pin down (i.e. exemplify) different concepts
+the formal expression of refinement, enables to pin down (i.e. exemplify) different concepts
 of refinement (data refinement, action refinement, trace refinement (aka spec refinement) and
@@
 
-\subsection{Relevance of the approach}
+\subsection{Summary and discussion of relevance of the approach}
+We can establish in our formal framework an attack that even a system using changing
+Ephemeral ids can be broken if the attacker physically follows a victim. This is a basic attack on
+anonymity: a user's connection between his Ephemeral Ids and personal details (Iphone MAC or name)
+is revealed to the attacker. The protection goal of privacy is thereby destroyed.
+
+When establishing the attack we start from a simplistic scenario that does not use Ephemeral Ids but fixed ids.
+In this (over)simplified model the attack is established. We then define a formal Refinement calculus for
+the Isabelle Infrastructure framework to refine the attack to a system with Ephemeral Ids that change in fixed
+time intervals obfuscating the relationship between user and his pseudonym\footnote{We identify the smartphone
+  and the user which might be also recognized by his appearance (face)}.
+
+Now, the refinement shows that although the Ephemeral Ids change regularly the same attack that has been
+identified on the very abstract level (fixed ids) persists.
+The refinement allows refining the datatype (Id $\mapsto$ EphId) but also delivers the usual trace
+refinement (behaviours of the refined systems are a subset of the traces of the abstract system).
+This persistence of the attack precisely shows which part of the system behaviour is responsible for the
+attack. In a second refinement step using action refinement based on this insight from the (repeated)
+attack, we can exclude the dangerous attack trace\footnote{Implication on the actual protection goals
+  that are endangered by the attack on privacy are discussed in the larger context of the Corona-warning
+  app (see previous Section)}.
+
+Clearly, the abstract attack we establish is obvious on an informal level but the persistence of the attack
+on the refined system is less obvious. The remedy by a second refinement step is an evident restriction of
+the system behaviour which gives a clear specification of a system secured against this attack. The use of
+formal methods therefore lies not in the discovery of an obvious attack on a simplified system but in showing
+how a formal specification including security refinement can lead to a stepwise improvement that is accompanied
+by formal proof in the Isabelle Infrastructure framework. The solution to exclude the attack in the second refinement
+step binds the action move together with a put action. %which effectively represents an action refinement.
+This shows that besides datatype refinement and trace refinement our Refinement calculus also entails action
+refinement. This action refinement is implemented implicitly by changing the effects of the actions
+in the semantic state transition relation. In future work, we could think about making the action refinement
+more explicit by considering a relationship between semantic rules or by refining the refinement notion to
+a more explicit layer of protocol steps -- similar to what has been done in previous applications of the
+Isabelle Infrastructure framework for example to Auction protocols \cite{kkp:17} or the Quantum Key Distribution \cite{kam:19c}.
+
+The security refinement in general might seem pointless, as in the first step of refinement the attack
+persists and even though the second refinement gets rid of this specific attack, it doesn't exclude the
+reachability of the attack goal altogether (if the attacker Eve gets Bob on his own in a location she can
+map all used EphIDs to him).
+However, the refinement makes the system relatively more secure in that for a larger number of traces the abstract
+attack does not work anymore\footnote{Adding probabilities as in \cite{kam:19c} enables quantifying this.}. 
+It is important to emphasize that security refinement is a cyclic process that
+improves security but does not usually terminate (like a loop) with a fixed point of 100\% secure system.
+In a refined model new detail may give rise to new attack possibilities. These additional attacks can be
+identified using the Attack Tree calculus and trigger further refinement steps.
 

### 2020-07-09 paper.tex
I've added the sections Model and Analysis (making them one section).

@@
 \newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
+\newcommand\ttbigcap{\mbox{{$\bigcap$}}}
 \newcommand\tttau{\mbox{{$\tau$}}}
@@
 \newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
+\newcommand\ttrefV{\mbox{{$\sqsubseteq_V$}}}
+\newcommand{\ttcalN}[1]{\mbox{{${\mathcal{N}}_{\texttt{#1}}$}}} 
 \newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
@@
 \cite{suc:16} and framework \cite{kam:19a}.
-
-\subsubsection{Overview}
-\label{sec:isainfover}
+%\subsubsection{Overview}
+%\label{sec:isainfover}
 Figure \ref{fig:theorystruc} gives an overview of the Isabelle Infrastructure 
@@
 \end{figure}
+
+\subsubsection{Kripke structures, CTL, and Attack Trees}
+\label{sec:intro}
+The Isabelle framework has now after various case studies become a general framework for
+the state-based security analysis of infrastructures with policies and actors. Temporal logic
+and Kripke structures build the foundation. 
+Meta-theoretical results have been established to show equivalence between attack trees 
+and CTL statements \cite{kam:18b}. 
+%As part of the meta-theory Correctness and Completeness have been 
+%proved in Isabelle \cite{kam:18b} and can be used to navigate between CTL and attack 
+%trees to find attacks. 
+This foundation provides a generic notion of state transition on which attack trees and
+temporal logic can be used to express properties. 
+The main notions used in this paper are:
+\begin{itemize}
+\item {Kripke structures and state transitions:}\\ 
+Using a generic state transition relation $\mapsto$, Kripke structures
+are defined as a set of states reachable by $\mapsto$ from an initial
+state set, for example
+\begin{ttbox}
+Kripke \{t. \ttexists i \ttin I. i \ttrelIstar t\} I
+\end{ttbox} 
+\item {CTL statements:}\\ 
+For example, we can write 
+\begin{ttbox}
+K \ttvdash {\sf EF} s
+\end{ttbox}
+to express that in Kripke structure \texttt{K} there is a path on which
+the property \texttt{s} (a set of states) will eventually hold.
+\item {Attack trees:} \\
+The datatype of attack trees has three constructors: 
+$\oplus_\vee$ creates or-trees and $\oplus_\wedge$ creates 
+and-trees.
+And-attack trees $l \ttattand s$ and or-attack trees $l \ttattor s$ 
+consist of a list of sub-attacks -- again attack trees. The third constructor 
+creates a base attack as a pair of state sets written \texttt{\ttcalN{(I,s)}}.
+For example, a two step and-attack leading from state set \texttt{I} via
+\texttt{si} to \texttt{s} is expressed as
+\begin{ttbox}
+\ttvdash [\ttcalN{(I,si)},\ttcalN{(si,s)}]\ttattand{\texttt{(I,s)}}
+\end{ttbox}
+%\begin{ttbox}
+%\ttvdash [\ttcalN{(I,s)},\ttcalN{(s,s')}]\ttattand{(I,s')}
+%\end{ttbox}
+\item {Attack tree refinement, validity and adequacy:}\\
+Attack trees have their own refinement (not to be mixed up with the
+system refinement presented in this paper as introduced in the next section).
+An abstract attack tree may be refined by spelling out the attack steps until a valid attack
+is reached:
+
+\texttt{\ttvdash A :: (\ttsigma :: state) attree)}.
+
+The validity is defined constructively (code is generated from it)
+and its adequacy with respect to a formal semantics in CTL is proved
+and can be used to facilitate actual application verification as demonstrated 
+her in the stepwise system refinements.
+\end{itemize}
 
@@
 
-\section{Model}
+\section{Model for Corona-warning app}
 \label{sec:model}
 
-\section{Analysis}
+\subsection{Infrastructures, Policies, and Actors}
+\label{sec:infra}
+% Adapted from CSF paper
+The Isabelle Infrastructure framework supports the representation of infrastructures 
+as graphs with actors and policies attached to nodes. These infrastructures 
+are the {\it states} of the Kripke structure. % for the attack trees.
+
+The transition between states is triggered by non-parametrized
+actions \texttt{get}, \texttt{move}, and \texttt{put} 
+executed by actors. 
+Actors are given by an abstract type \texttt{actor} and a function 
+\texttt{Actor} that creates elements of that type from identities 
+(of type \texttt{string} written \texttt{''s''} in Isabelle). 
+Actors are contained in an infrastructure graph type \texttt{igraph}
+constructed by its constructor \texttt{Lgraph}. 
+\begin{ttbox}
+{\bf datatype} igraph = 
+         Lgraph (location \tttimes location)set 
+                 location \ttfun identity set
+                 identity \ttfun (string set \tttimes string set \tttimes efid)  
+                 location \ttfun string \tttimes (dlm \tttimes data) set
+                 location \ttfun efid set
+                 actor \ttfun location \ttfun (identity \tttimes efid) set
+
+\end{ttbox}
+In the current application of the framework to the Corona-warning app case study, 
+this graph contains a set of location pairs representing the %structure 
+topology of the infrastructure
+as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
+%e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
+%function returning True for any input $x$.} 
+that assigns a set of actor identities to each node (location) in the graph.
+The third component of an \texttt{igraph} assigns the credentials to each actor: 
+a triple-valued function whose first range component is a set describing the credentials 
+in the possession of an actor and the second component 
+is a set defining the roles the actor can take on; 
+most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
+just a natural number but will be refined to actually represent lists of ``Ephemeral'' Ids later
+when refining the specification.
+\begin{ttbox}
+{\bf datatype} efid = Efid nat
+\end{ttbox}
+The fourth component of the type \texttt{igraph}
+assigns security labelled data to locations, a feature not used in the
+current application.
+The second to last component assigns the set of efids of all currently present smart phones
+to each location of the graph. The last component finally denotes the knowledge set of each
+actor for each location: a set of pairs of actors and potential ids.
+
+Corresponding projection functions for each of the components of an 
+infrastructure graph are provided; they are named \texttt{gra} for the actual 
+set of pairs of locations, \texttt{agra} for the actor map, \texttt{cgra} for 
+the credentials, \texttt{lgra} for the %state of a location and the 
+data at that location, \texttt{egra} for the assignment of current efids to locations,
+and \texttt{kgra} for the knowledge set for each actor for each location.
+
+In the Corona-virus warning app, the initial values for the \texttt{igraph} components use
+two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
+data map component \texttt{ex\_locs}).
+\begin{ttbox}
+ex_loc_ass \ttequiv (\ttlam x. if x = pub then \{''Alice'', ''Bob'', ''Eve''\}  
+                 else (if x = shop then \{''Charly'', ''David''\} 
+                 else \{\}))
+ex_creds \ttequiv (\ttlam x. if x = ''Alice'' then (\{\}, \{\}, Efid 1) else 
+              (if x = ''Bob'' then  (\{\},\{\}, Efid 2) else 
+              (if x = ''Charly'' then (\{\},\{\}, Efid 3) else
+              (if x = ''David'' then (\{\},\{\}, Efid 4) else
+              (if x = ''Eve'' then (\{\},\{\}, Efid 5)
+               else (\{\},\{\},Efid 0))))))
+ex_efids \ttequiv (\ttlam x. if x = pub then \{Efid 1, Efid 2, Efid 5\}
+                  else (if x = shop then \{Efid 3, Efid 4\} else \{\}))
+ex_knos \ttequiv (\ttlam x. (if x = Actor ''Eve'' then (\ttlam l. \{\} else (\ttlam l. \{\})))
+\end{ttbox}  
+These components are wrapped up into the following \texttt{igraph}.
+\begin{ttbox}
+ex_graph \ttequiv
+      Lgraph \{(pub, shop)\} ex_loc_ass ex_creds ex_locs ex_efids ex_knos
+\end{ttbox}  
+Infrastructures are given by the following datatype that 
+contains an infrastructure graph of type \texttt{igraph} 
+and a policy given by a function that assigns local policies over a graph to
+all locations of the graph. 
+\begin{ttbox}
+{\bf{datatype}} infrastructure = Infrastructure igraph 
+                                         [igraph, location] \ttfun policy set
+\end{ttbox}
+There are projection functions \texttt{graphI} and \texttt{delta} when applied
+to an infrastructure return the graph and the local policies, respectively.
+The function \texttt{local\_policies} gives the policy for each location \texttt{x}
+over an infrastructure graph \texttt{G} as a pair: the first element of this pair is 
+a function specifying the actors \texttt{y} that are entitled to perform the actions 
+specified in the set which is the second element of that pair.
+The local policies definition for the Corona-warning app, simply permits all actions to
+all actors in both locations.
+\begin{ttbox}
+  local_policies G \ttequiv
+         (\ttlam x. if x = pub then  \{(\ttlam y. True, \{get,move,put\})\}
+          else (if x = shop then \{(\ttlam y. True, \{get,move,put\})\} else \{\}))
+\end{ttbox}  
+For the Corona-warning app, the initial infrastructure contains the graph \texttt{ex\_graph}
+with its two locations pub and shop and is then wrapped up with the local policies
+into the infrastructure \texttt{Corona\_scenario} 
+that represents the ``initial'' state for the Kripke structure.
+\begin{ttbox}
+ Corona_scenario \ttequiv Infrastructure  ex_graph local_policies
+\end{ttbox}
+
+\subsection{Policies, privacy, and behaviour}
+\label{sec:}
+Policies specify the expected behaviour of actors of an infrastructure. 
+They are given by pairs of predicates (conditions) and sets of (enabled) actions.
+%\begin{ttbox}
+%{\bf type}_{\bf{synonym}} policy = ((actor \ttfun bool) \tttimes action set)
+%\end{ttbox}
+%The behaviour of actors is 
+They are defined by the \texttt{enables} predicate:
+an actor \texttt{h} is enabled to perform an action \texttt{a} 
+in infrastructure \texttt{I}, at location \texttt{l}
+if there exists a pair \texttt{(p,e)} in the local policy of \texttt{l}
+(\texttt{delta I l} projects to the local policy) such that the action 
+\texttt{a} is a member of the action set \texttt{e} and the policy 
+predicate \texttt{p} holds for actor \texttt{h}.
+\begin{ttbox}
+enables I l h a \ttequiv \ttexists (p,e) \ttin delta I l. a \ttin e \ttand p h
+\end{ttbox}
+
+The Privacy protection goal is to avoid deanonymisation. That is, an attacker should not be able to
+disambiguate the set of pairs of real Ids and EphIds. This is abstractly expressed in the predicate
+identifiable.
+\begin{ttbox}
+identifiable eid A \ttequiv is_singleton\{(Id,Eid). (Id, Eid) \ttin A \ttand Eid = eid\}
+\end{ttbox}
+The predicate identifiable is used to express as the global policy `Eve cannot deanonymize an Ephemeral
+Id \texttt{eid} using the gathered knowledge':
+%{\bf fixes} global_policy::[infrastructure, identity] \ttfun bool
+%{\bf defines}  
+\begin{ttbox}
+  global_policy I eid \ttequiv
+             \ttneg(identifiable eid 
+                ((\ttbigcap (kgra(graphI I)(Actor ''Eve'')`(nodes(graphI I))))
+                 - \{(x,y). x = ''Eve''\}))
+\end{ttbox}
+
+\subsection{Infrastructure state transition}
+\label{sec:statetrans}
+The state transition relation uses the syntactic infix notation 
+\texttt{I \ttrelI\, I'}  to denote that infrastructures 
+\texttt{I} and \texttt{I'} are in this relation.
+%Rules for the put, get, move, process and delete actions exist.
+To give an impression of this definition, we show first the rule defining the state transition for
+the action get. Initially, this rule expresses that
+an actor that resides at a location \texttt{l} (\texttt{a \ttatI\ l})
+and is enabled by the local policy in this location to ``get'' can
+combine  all ids at the current location (contained in \texttt{egra G l}) with
+all actors at the current location (contained in \texttt{agra G l}) and add this
+set of pairs to his knowledge set \texttt{kgra G} using the function update \texttt{f(l := n)}
+redefining the function \texttt{f} for the input \texttt{l} to have now the new value \texttt{n}.
+\begin{ttbox}
+ {\bf{get\_data}}: G = graphI I \ttImp a \ttatI l \ttImp enables I l (Actor a) get \ttImp 
+   I' = Infrastructure 
+          (Lgraph (gra G)(agra G)(cgra G)(lgra G)(egra G)
+            ((kgra G)((Actor a) := ((kgra G (Actor a))(l:=
+                 \{(x,y). x \ttin agra G l \ttand y \ttin egra G l\})))))
+          (delta I)
+   \ttImp I \ttrel{n} I' 
+\end{ttbox}
+Another interesting rule for the state transition is the one for \texttt{move} whose structure
+resembles the previous one.
+\begin{ttbox}
+ {\bf{move}}: G = graphI I \ttImp  a \ttin actors_graph(graphI I) \ttImp
+    l \ttin nodes G \ttImp l' \ttin nodes G \ttImp enables I l' (Actor h) move \ttImp
+    I' = Infrastructure (move_graph_a a l l' (graphI I))(delta I)
+ \ttImp I \ttrelI I' 
+\end{ttbox}
+The semantics of this rule is embedded in the function \texttt{move\_graph\_a}
+that adapts the infrastructure state so that the moving actor \texttt{a} is now
+associated to the target location \texttt{l'} in the actor map \texttt{agra} and not
+any more at \texttt{l} and also the association of \texttt{efids} is updated accordingly.
+\begin{ttbox}
+move_graph_a n l l' g \ttequiv
+  Lgraph (gra g) 
+         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then 
+           ((agra g)(l := (agra g l) - \{n\}))(l' := (insert n (agra g l')))
+          else (agra g))
+         (cgra g)(lgra g)
+         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then
+               ((egra g)(l := (egra g l) - \{efemid (cgra g n)\}))
+                        (l' := (insert (efemid (cgra g n))(egra g l')))
+          else egra g)
+         (kgra g)
+\end{ttbox}  
+Based on this state transition and the above defined \texttt{Corona\_scenario}, we define the first Kripke structure.
+\begin{ttbox}
+ corona_Kripke \ttequiv Kripke \{ I. Corona_scenario \ttrelIstar I \} \{Corona_scenario\}
+\end{ttbox}
+
+\subsection{Attack analysis}
 \label{sec:ana}
+For the analysis of attacks, we negate the security property that we want to achieve,
+usually the global policy.
+
+Since we consider a predicate transformer semantics, we use
+sets of states to represent properties. 
+%For example, the attack property
+The invalidated global policy is given by the following set \texttt{shc}.
+\begin{ttbox}
+ scorona \ttequiv \{x. \ttexists n. \ttneg global_policy x (Efid n)\}
+\end{ttbox}
+The attack we are interested in is to see whether for the scenario
+\begin{ttbox}
+ Kripke\_scenario \ttequiv  Infrastructure ex_graph local_policies 
+\end{ttbox}
+from the initial state \texttt{Icorona \ttequiv \{corona\_scenario\}},
+the critical state \texttt{scorona} can be reached,
+that is, is there a valid attack \texttt{(Icorona,scorona)}?
+
+For the Kripke structure \texttt{corona\_Kripke}
+we first derive a valid and-attack using the attack tree proof calculus.
+\begin{ttbox}
+  \ttvdash [\ttcalN{(Icorona,Corona)},\ttcalN{(Corona,Corona1)},\ttcalN{(Corona1,Corona2)},
+             \ttcalN{(Corona2,Corona3)},\ttcalN{(Corona3,scorona)}]\ttattand{\texttt{(Icorona,scorona)}}
+\end{ttbox}
+The sets \texttt{Corona, Corona1, Corona2, Corona3} are the intermediate states where
+\texttt{Bob} moves to shop and \texttt{Eve} follows him collecting the Ephemeral Ids in each location.
+The collected information enables to identify Bob's Ephemeral Id.
+
+The attack tree calculus \cite{kam:18b} exhibits that an attack is possible.
+\begin{ttbox}
+ corona_Kripke \ttvdash {\sf EF} scorona
+\end{ttbox}
+We can simply apply the Correctness theorem \texttt{AT\_EF} to 
+immediately prove this CTL statement. This application of the meta-theorem 
+of Correctness of attack trees saves us proving the CTL formula tediously 
+by exploring the state space in Isabelle proofs. Alternatively, we could use 
+the generated code for the function \texttt{is\_attack\_tree} in Scala 
+(see Section \ref{sec:at}) to check that a refined attack of the above is valid.
 
@@
 
-\section{Refining the Specification}
+\subsection{Refining the Specification}
 \label{sec:corref}

### 2020-07-09 paper.tex
renamed ModTrans inot Refeinemnt

@@
 the generated code for the function \texttt{is\_attack\_tree} in Scala 
-(see Section \ref{sec:at}) to check that a refined attack of the above is valid.
+(see \cite{kam:18b}) to check that a refined attack of the above is valid.
 

### 2020-07-09 paper.tex
small edits and added put into Infrastructure

@@
 \newcommand\ttecal{\mbox{$\mathcal{E}$}}
-\newcommand\ttimg{\mbox{$\triangleleft$}}
+%\newcommand\ttimg{\mbox{$\triangleleft$}}
+\newcommand\ttimg{\mbox{\texttt{`}}}
 
@@
 \begin{ttbox}
- mod_trans ::  (\ttsigma Kripke \tttimes (\ttsigma' \ttfun \ttsigma) \tttimes \ttsigma' Kripke) 
-               \ttfun bool                  ("_ \ttmref{(\_)} _")
+ refinement :: (\ttsigma Kripke \tttimes (\ttsigma' \ttfun \ttsigma) \tttimes \ttsigma' Kripke) \ttfun bool ("_ \ttmref{(\_)} _")
   K \ttmeref K' \ttequiv \ttforall s' \ttin states K'. \ttforall s \ttin init K'. 
-             s \ttrelstar{\sigma'} s' \ttimp \ttecal(s) \ttin init K 
-             \ttand \ttecal(s) \ttrelstar{\sigma} \ttecal(s')
+             s \ttrelstar{\sigma'} s' \ttimp \ttecal(s) \ttin init K \ttand \ttecal(s) \ttrelstar{\sigma} \ttecal(s')
 \end{ttbox}
-The definition of \texttt{K \ttmeref\, K'} states that for any state $s$ 
+The definition of \texttt{K \ttmeref\, K'} states that for any state $s'$ 
 of the refined Kripke structure that can be reached by the state transition
-in zero or more steps from an initial state $s_0$ of the refined Kripke 
+in zero or more steps from an initial state $s$ of the refined Kripke 
 structure, the mapping ${\mathcal E}$ from the refined to the abstract 
 model's state must preserve this reachability, i.e., the image of
-$s_0$ must also be an initial state and from there the image of $s$
+$s$ must also be an initial state and from there the image of $s'$
 under ${\mathcal E}$ must be reached with $0$ or $n$ steps.
@@
    K \ttmeref K'  \ttImp init K \ttsubseteq \ttecal\ttimg(init K') \ttImp
-   \ttforall s' \ttin Pow(states K'). K' \ttvdash {\sf EF} s' 
-              \ttimp K \ttvdash {\sf EF} (\ttecal\ttimg(s'))
+   \ttforall s' \ttin Pow(states K'). K' \ttvdash {\sf EF} s' \ttimp K \ttvdash {\sf EF} (\ttecal\ttimg(s'))
 \end{ttbox}

### 2020-07-10 paper.tex
started writing section on Refinement application

@@
 
-\section{Model for Corona-warning app}
+\section{Modeling and analyzing Corona-warning app}
 \label{sec:model}
-
 \subsection{Infrastructures, Policies, and Actors}
@@
                  actor \ttfun location \ttfun (identity \tttimes efid) set
-
 \end{ttbox}
@@
 action refinement, traces may be refined by combining consecutive system events
-into atomic events thereby reducing traces. This will be detailed on the example
-below.
+into atomic events thereby reducing traces. This will be detailed in the second refinement step
+of the Corona-virus warning app considered next.
 
@@
 \label{sec:corref}
+Clearly, fixed Ephemeral Ids are not really ephemeral. The model presented
+in Section \ref{sec:model} is deliberately been chosen very abstractly to allow focusing on
+the basic system architecture and finding an initial deanonymization attack.
+We now introduce ``proper'' Ephemeral Ids and show how the system datatype can be refined to a system
+that uses those instead of the fixed ones.
+
+For the DP-3T Ephemeral Ids \TODOref{DP-3T Whitepaper}, for each day $t$ a seed $SK_t$
+is used to generate a list of length $n = 24 * 60 / L$, where $L$ is the duration for which
+the Ephemeral Ids are posted by the smart phone
+\begin{ttbox}
+  EphId1 || ... || EphIdn = PRG(PRF(SKt,``broadcast key''))
+\end{ttbox} 
+``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
+fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \TODOref{DP-3T Whitepaper}.
+
+From a cryptographic point of view, the crucial properties of the Ephemeral Ids are that
+they are purely random, therefore, they cannot be guessed, but at the same time if -- after
+the actual encounter between sender and receiver -- the seed $SK_t$ is published, it is feasible
+to relate any of the \texttt{EphIDi} to $SK_t$  for all \texttt{i} $\in \{1, \ldots, n\}$.
+For a formalisation of this crucial cryptographic property in Isabelle it suffices to define a new type
+of list of Ephemeral Ids \texttt{efidlist} containing the root $SK_t$ (the first \texttt{efid}),
+a current \texttt{efid} indicated by a list pointer of type \texttt{nat}, and the actual list of
+\texttt{efid}s.
+\begin{ttbox}
+{\bf{datatype}} efidlist = Efids "efid" "nat" "efid list"
+\end{ttbox}
+We define functions for this datatype: \texttt{efidsroot} returning the first of the three
+consituents in an \texttt{efidlist} (the root $SK_t$); \texttt{efids\_index} giving the second
+component, the index of the current \texttt{efid}; \texttt{efids\_inc\_ind} applied to an
+\texttt{efidlist} increments the index; \texttt{efids\_cur} returning the current \texttt{efid}
+from the list and \texttt{efids\_list} for the entire list (the third component).
 

### 2020-07-10 paper.tex
finished a first draft of Sectron 4.2 Refining the specification

@@
 \begin{ttbox}
- {\bf{move}}: G = graphI I \ttImp  a \ttin actors_graph(graphI I) \ttImp
+ {\bf{move}}: G = graphI I \ttImp a \ttatI l \ttImp a \ttin actors_graph(graphI I) \ttImp
     l \ttin nodes G \ttImp l' \ttin nodes G \ttImp enables I l' (Actor h) move \ttImp
@@
 %For example, the attack property
-The invalidated global policy is given by the following set \texttt{shc}.
+The invalidated global policy is given by the set \texttt{scorona}.
 \begin{ttbox}
@@
 \texttt{Bob} moves to shop and \texttt{Eve} follows him collecting the Ephemeral Ids in each location.
-The collected information enables to identify Bob's Ephemeral Id.
+The collected information enables identifying Bob's Ephemeral Id.
 
@@
 Clearly, fixed Ephemeral Ids are not really ephemeral. The model presented
-in Section \ref{sec:model} is deliberately been chosen very abstractly to allow focusing on
+in Section \ref{sec:model} has deliberately been designed abstractly to allow focusing on
 the basic system architecture and finding an initial deanonymization attack.
@@
 from the list and \texttt{efids\_list} for the entire list (the third component).
+
+The first step of refinement replaces the simple \texttt{efid} in the infrastructure type by the
+new type \texttt{efidlist}. Note, that in the new datatype \texttt{igraph} this change affects only
+the thrird component, the credentials. The last two componenets, the set of currently present efids,
+and the knowledge set, still operate on the simple type \texttt{efid}.
+\begin{ttbox}
+{\bf datatype} igraph = 
+         Lgraph (location \tttimes location)set 
+                 location \ttfun identity set
+                 identity \ttfun (string set \tttimes string set \tttimes efidlist)  
+                 location \ttfun string \tttimes (dlm \tttimes data) set
+                 location \ttfun efid set
+                 actor \ttfun location \ttfun (identity \tttimes efid) set
+\end{ttbox}
+The refined state transition relation implements the possibility of changing the Ephemeral Ids
+by the rule for the action \texttt{put}.
+\begin{ttbox}
+ {\bf{put}}: G = graphI I \ttImp  a \ttatI l \ttImp enables I l (Actor a) put \ttImp
+      I' = Infrastructure (put_graph_efid A l (graphI I))(delta I)
+ \ttImp I \ttrelI I' 
+\end{ttbox}
+The change to the infrastructure state is implemented in the function \texttt{put\_graph\_efid} that
+increases the current index in the \texttt{efidlist} in the credential component \texttt{cgra g n} for
+the ``putting'' actor identity \texttt{n} and inserts the current \texttt{efid} from that credential
+component into the \texttt{egra} component, the set of currently ``present'' Ephemeral Ids at the location
+\texttt{l}.
+\begin{ttbox}
+put_graph_efid n l g  \ttequiv
+  Lgraph (gra g)(agra g)
+         ((cgra g)(n := (credentials (cgra g n), roles (cgra g n),
+                         efids_inc_ind(efemid (cgra g n)))))
+         (lgra g)
+         ((egra g)(l := insert (efids_cur(efemid (cgra g n)))(egra g l)))
+         (kgra g)
+\end{ttbox}
+We can now apply the refinement by defining a datatype map from the refined infrastructure type
+\texttt{InfrastructureOne.infrastructure} to the former one \texttt{Infrastructure.infrastructure}.
+\begin{ttbox}
+{\bf{definition}} refmap :: InfrastructureOne.infrastructure \ttfun
+                          Infrastructure.infrastructure
+{\bf{where}} refmap I =
+  Infrastructure.Infrastructure 
+    (Infrastructure.Lgraph
+       (InfrastructureOne.gra (graphI I))
+       (InfrastructureOne.agra (graphI I))
+        (\ttlam h. repl_efr 
+           (InfrastructureOne.cgra (graphI I)) h)
+           (InfrastructureOne.lgra (graphI I))
+           (InfrastructureOne.egra (graphI I))
+           (\ttlam a. \ttlam l.
+  (\ttlam (x,y).(x, efids_root(efemid(InfrastructureOne.cgra (graphI I) x))))
+                       \ttimg(InfrastructureOne.kgra (graphI I)) a l))                                                    
+\end{ttbox}  
+This is then plugged into the parameter \texttt{\ttecal} of the refinement operator allowing to
+prove \texttt{corona\_Kripke \ttref{\texttt{refmap}} corona\_KripkeO} where the latter is the
+refined Kripke structure.
+
+Surprisingly, we can still prove \texttt{corona\_KripkeO \ttvdash EF scoronaO} by using the
+same attack tree as in the abstract model: if Bob moves from pub to shop, he is vulnerable to
+being identifiable as long as he does not change the current \texttt{efid}. So, if Eve moves to
+the shop as well and performs a get before Bob does a put, then Eve's knowledge sets permits identifying
+Bob's current Ephemeral Id as his.
+
+\subsubsection{Second refinement step}
+The persistent attack can be abbreviated informally by the action sequence \texttt{[get,move,move,get]}
+performed by actors Eve, Bob, Eve, and Eve again, respectively.
+How can a second refinement step avoid that Eve does the last \texttt{get} by imposing that after the
+first \texttt{move} of Bob a \texttt{put} action must happen before Eve can do another \texttt{get}?
+A very simple remedy to exclude this attack is to bind a \texttt{put} action after every \texttt{move}.
+We can implement that change by a minimal update to the function \texttt{move\_graph\_a} (see Section \ref{sec:model})
+by adding an increment (highlighted in the code snippet)
+of the currently used Ephemeral Id before updating the \texttt{egre} component of the target location.
+\begin{ttbox}
+move\_graph\_a n l l' g \ttequiv ...
+  (l' := insert (efids_cur({\bf efids\_inc\_ind}(efemid (cgra g n))))(egra g l))...
+\end{ttbox}
 

### 2020-07-10 paper.tex
a few more lines to complete the Refinement section

@@
 \end{ttbox}
-
 This idea of property preservation coincides with the classical idea of
@@
 
-In addition, as we can observe in the following example in the second refinement
-step, we also have a simple implicit version of {\it action refinement}. In an
+In addition, we also have a simple implicit version of {\it action refinement}. In an
 action refinement, traces may be refined by combining consecutive system events
-into atomic events thereby reducing traces. This will be detailed in the second refinement step
+into atomic events thereby reducing traces.
+We can observe this kind of refinement in the second refinement step
+%This will be detailed in the second refinement step
 of the Corona-virus warning app considered next.
@@
 \end{ttbox}  
+This is an action refinement because the move action is changed. It is a refinement, since
+any trace of the refined model can still be mapped to a trace in the more abstract model just
+omitting a few steps (the refinement relation is defined using the reflexive transitive closure
+\texttt{\ttrelIstar}.
 

### 2020-07-10 paper.tex
Introduction and Related works (incl. graphic)

@@
 \begin{abstract}
-We provide a model in the Isabelle Infrastructure framework of the recently published
+We provide a model in the Isabelle Infrastructure framework of the recently published German
 Corona-virus warning app. The app supports breaking infection chains by informing users
@@
 The German Chancellor Angela Merkel has strongly supported the publication of
-the mobile phone Corona warning app by publicly proclaiming that the ``Corona
+the mobile phone Corona warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
 App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
 in Germany have downloaded the app with 6 million on the first day.
-This app is one amongst many similar application that aim at the very important goal
-to ``break infection chains'' by providing timely information of users whether they
-have been exposed to close contact with a person that has been infected.
+CWA is one amongst many similar applications that aim at the very important goal
+to ``break infection chains'' by providing timely information to users of whether they
+have been in close proximity to a person who tested positive for COVID-19.
 
-The Corona-virus warning app
+%The Corona-virus warning app
 %has taken a long time to develop being published only on
 %16th June 2020. It
-was a quite costly project but this was mainly due to the management of
-Telekom and SAP being in the driving seat. But the app has been designed with great
-attention on privacy:  a distributed architecture \cite{} has been adopted after a long and
+%was a quite costly project but this was mainly due to the management of
+%Telekom and SAP being in the driving seat. But the app has been designed with great
+The app was designed with great
+attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted after a long and
 heated debate with supporters of a central architecture. The distributed architecture is
-based on a very clever distributed application design whereby users' phones are sending
-highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
-protocol. The app saves those IDs of people in close proximity. When at a later date an
-infected person reports his infection to a central server, the unique root ID is published
-and in the daily check all mobile phones connecting to the central server download
-the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
-all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
-regularly check whether their user has been exposed to an infected person and issue
-a warning to the user and recommendation to contact health authorities.
+based on a very clever application design whereby users' phones broadcast
+highly anonymized so called ``Ephemeral IDs'' at physical locations via the
+Bluetooth Low Energy Beacon protocol.
+The app saves those IDs of people in close proximity.
+When at a later date an infected person reports his infection to a central server,
+all clients are provided the means to reconstruct the IDs used by this particular
+person's device over the last 14 days.
+Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
+%
+%, the unique root ID is published
+%and in the daily check all mobile phones connecting to the central server download
+%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
+%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
+%regularly check whether their user has been exposed to an infected person and issue
+%a warning to the user and recommendation to contact health authorities.
 %The warning issued by the Corona warning app entitles to
@@
 
-The Corona-virus app has been produced based on a sophisticated security
-concept conceived by experts in the field, %to our knowledge, no formal verification
+The app now in use in Germany has been developed based on a protocol proposed by the DP-3T\cite{dp3t:github} project:
+A sophisticated security concept
+conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
+%The Corona-virus app has been produced based on a sophisticated security
+%concept conceived by experts in the field %to our knowledge, no formal verification
+%%
+%and has strong claims with regard to mathematical support:
 %
-and has strong claims with regard to mathematical support:
-
-``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
+%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
 %  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
+%
 However, there has been, as of yet, to our knowledge no formal verification been involved.
@@
 it allows to reveal weak points of the architecture, show that the measures that have
-been conceived are suitable to cover those weak points. Thereby, we believe
-that our current work is useful to increase the trust in the Corona-virus app necessary for
-its wide adoption which in turn is crucial for it to be efficient.
+been conceived are suitable to cover those weak points,
+or where and to what extend trade offs have to be made due to inherent vulnerabilities.
+Given that the protocol implemented by the framework CWA runs on resembles only the most basic
+of the three proposed by the DP-3T project,
+a formal verified model that stresses the impact or limits of certain security measures
+might help lend some more weight to appeals like \cite{dp3t:impl}
+%to Apple and Google
+to adopt the more sophisticated protocols.
+%
+%Thereby, we believe
+%that our current work is useful to increase the trust in the Corona-virus app necessary for
+%its wide adoption which in turn is crucial for it to be efficient.
 
@@
 \label{sec:history}
-This paper is mainly concerned with the architecture and protocols proposed by the
-Decentralized Privacy-Preserving Proximity Tracing project (DP-3T).
+We are mainly concerned with the architecture and protocols proposed by the
+DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project.
 The main reason to focus on this particular family of protocols is the
-\textit{Exposure Notification Framework}, jointly published by Apple and Google, that adopts
-the main core principles of the DP-3T proposal. This API is not only used in the German
+\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google, that adopts
+%the main
+core principles of the DP-3T proposal. This API is not only used in the German
 Corona-virus warning app but has the potential of being widely adopted in future
@@
 DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
-  As we will see, DP-3T involves a central backend server. It is decentralized with regard to
-  collection and evaluation of contact information:
+  DP-3T involves a central backend server. It is decentralized with regard to
+  the collection and evaluation of contact information:
   In centralized architectures the server provides a risk scoring services, whereas decentralized
@@
   the backend.}
+\\
 
-\subsubsection{Basic DP-3T protocol: How does it work?}
+All DP-3T protocols roughly work the same:
+%\subsubsection{Basic DP-3T protocol: How does it work?}
 %  How does it relate to ENF/CWA? (broad strokes)}
 %
-Upon installation, the app generates secret daily seeds to derive so called \textit{Ephemeral IDs}
+Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral IDs}
 (EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
@@
 with associated meta-data such as signal attenuation and date. In DP-3T the contact information
-gathered is never shared but only evaluated locally (on the device).
+gathered is never shared but only evaluated locally.
 
-If patients are tested positive on the Corona-virus, they are entitled to upload specific data to a
-central backend server. This data is accumulated by the backend server and redistributed to all
+If patients test positive for the Corona-virus, they are entitled to upload specific data to a
+central backend server. This data is aggregated by the backend server and redistributed to all
 clients regularly to provide the means for local risk scoring, i.\,e., determining whether collected
@@
 this basically translates into publishing the daily seeds used to derive EphIDs from.
-Aside from additional security measures,
+The protocol implemented by ENF and,
+hence, CWA adopts this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
+privacy and data protection properties to different degrees but are more costly to implement.
+%Aside from additional security measures,
 %\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
-like signing content the server provides as a general rule, the protocol implemented by ENF and,
-hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
-privacy and data protection properties to different degrees but are more costly.
-%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
+%like signing content the server provides as a general rule, the protocol implemented by ENF and,
+%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
+%privacy and data protection properties to different degrees but are more costly.
+%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
+
+Figure \ref{fig:dp3tprot} illustrates the basic system architecture
+along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=5in]{DP-3T_basic_protocol}
+  \end{center}
+  \caption{Decentralized privacy-preserving proximity tracing protocol
+    (as implemented by CWA)}
+  \label{fig:dp3tprot}
+\end{figure}
 

### 2020-07-10 paper.tex
insider.bib completed and most of the TODOs fixed

@@
 
-The app now in use in Germany has been developed based on a protocol proposed by the DP-3T\cite{dp3t:github} project:
+The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
 A sophisticated security concept
@@
 We are mainly concerned with the architecture and protocols proposed by the
-DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project.
+DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project \cite{dp3t:github}.
 The main reason to focus on this particular family of protocols is the
@@
 There are, however, competing architectures noteworthy, namely protocols developed under the
-roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT), e.g.
-PEPP-PT-ROBERT \TODOref{ROBERT github repository} and PEPP-PT-NTK\TODOref{NTK ???; mentioned in
-  ``Response to Analysis of DP-3T''}.
+roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT) \cite{pepppt:github}, e.g.
+PEPP-PT-ROBERT \cite{pepppt:ROBERT}.
 
@@
 
-In the most simple (and insecure) protocol proposed by DP-3T
-\TODOref{Low-cost decentralized proximity tracing; DP-3T Whitepaper p. 14ff}
+In the most simple (and insecure) protocol proposed by DP-3T %\cite{dp3t:wp}[p14ff]
 this basically translates into publishing the daily seeds used to derive EphIDs from.
 The protocol implemented by ENF and,
-hence, CWA adopts this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
+hence, CWA adopts this low-cost design \cite{cwa:arch}. DP-3T proposes two other, more sophisticated protocols that improve
 privacy and data protection properties to different degrees but are more costly to implement.

### 2020-07-11 paper.tex
additional citations and "Protection goals" chapter (incl. graphic)

@@
 The main reason to focus on this particular family of protocols is the
-\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google, that adopts
+\textit{Exposure Notification Framework} (ENF) \cite{enf:proj}, jointly published by Apple and Google, that adopts
 %the main
@@
 %
-Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral IDs}
+Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
 (EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
@@
 \subsection{Protection goals of attacks}
+Bluetooth based proximity tracing gives rise to a variety of noteworthy privacy and security issues that deserve formal consideration (and verification).
+The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, connecting an Ephemeral Id to a particular device is the first step into multiple different attacks, we want to classify as deanonymisation attacks:
+\begin{itemize}
+\item \textbf{Tracking a device}
+  \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]\\
+  Bluetooth enabled devices introduce some fundamental risks, that are not
+  specific to proximity detection: If the Bluetooth signal is linked to a fixed
+  identifier, say a MAC address, constant broadcasting makes
+  a device trackable.
+  Periodically changing Ephemeral Ids can be an effective mitigation\footnote{%
+    In addition Bluetooth has to be configured to use rotating MAC addresses
+    instead of fixed ones, as is the default in modern smartphone operating
+    systems. Also, rotation of Bluetooth identifiers has to happen
+    in sync with MAC address randomization; the Apple/Google proposal
+    \cite{enf:bluetooth} ensures this.}
+  as they limit the time frame in which a certain id can be used to
+  track the corresponding device, hence, rendering the attack not impossible
+  but significantly less severe.
+\item \textbf{Identify infected individuals}
+  \cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]\\
+  Like any notification about, e.\,g., a STD, might inform a spouse about the
+  infidelity of the other, given the right circumstances, being notified about a
+  possible exposure to COVID-19 might reveal illicit information to
+  the recipient like who it was that tested positive for the virus.
+  This is an inherent vulnerability of any such system,
+  whether it is analogous or digital (digitally aided).\footnote{
+    As DP-3T emphasizes, it it, indeed, inherent to proximity tracing in general
+    \cite{dp3t:psre}[p6].
+    The likelihood of said circumstances to occur, however,
+    might increase when notifications are carried out automatically.
+    Therefore, we feel that it is, nonetheless, worth considering.}
+  In contrast to tracking a Bluetooth enabled device,
+  this attack is not considerably mitigated by changing Ephemeral Ids:
+  %Once connected to a device or individual, an Ephemeral Id might yield the
+  %desired information for up to 14 days
+  A positive test result can be reported for as much as 14 days after the
+  compromised Ephemeral Id was used and still reveal the desired information.
+\end{itemize}
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=6in]{DP-3T_deanonymisation_attacks}
+  \end{center}
+  \caption{Two deanynomisation attacks: Tracking users and identifying confirmed COVID-19 cases}
+  \label{fig:dp3tprot}
+\end{figure}
 

### 2020-07-11 paper.tex
techn. part: fixed TODOs (i.e. citations of DP-3T whitepaper)

@@
 \newcommand\rel{Re${\cal{L}}$}
-\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
-\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
-\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
+%\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
+%\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
+%\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
 \newcommand\ttand{\mbox{{$\land$}}}
@@
 
-For the DP-3T Ephemeral Ids \TODOref{DP-3T Whitepaper}, for each day $t$ a seed $SK_t$
+For the DP-3T Ephemeral Ids \cite{dp3t:wp}, for each day $t$ a seed $SK_t$
 is used to generate a list of length $n = 24 * 60 / L$, where $L$ is the duration for which
@@
 ``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
-fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \TODOref{DP-3T Whitepaper}.
+fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \cite{dp3t:wp}.
 

### 2020-07-11 paper.tex
paper, techn. part: minor fix (consistent spelling of Ephemeral Id)

@@
 by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
-to physical locations. The Ephemeral IDs, their sending and change is added to Infrastructures
+to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
 by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules

### 2020-07-11 paper.tex
shrunk the paper to 16 pages

--- 2020-07-11 paper.tex
@@
 \newcommand\rel{Re${\cal{L}}$}
-%\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
-%\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
-%\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
+\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
+\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
+\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
 \newcommand\ttand{\mbox{{$\land$}}}
@@
 The main reason to focus on this particular family of protocols is the
-\textit{Exposure Notification Framework} (ENF) \cite{enf:proj}, jointly published by Apple and Google, that adopts
+\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google, that adopts
 %the main
@@
 %
-Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
+Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral IDs}
 (EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
@@
 \begin{figure}[htb]
+\vspace{-.5cm}
   \begin{center}
@@
   \end{center}
-  \caption{Decentralized privacy-preserving proximity tracing protocol
-    (as implemented by CWA)}
+\vspace{-.5cm}
+  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
   \label{fig:dp3tprot}
+\vspace{-.5cm}
 \end{figure}
@@
 \begin{figure}
+\vspace{-.5cm}
   \begin{center}
-  \includegraphics[scale=.5]{theory_structure.pdf}
+  \includegraphics[scale=.35]{theory_structure.pdf}
 \end{center}
-%\vspace{-.5cm}
+\vspace{-.5cm}
 \caption{Generic Isabelle Infrastructure framework applied to Corona-warning app.}
 \label{fig:theorystruc}
+\vspace{-.5cm}
 \end{figure}
@@
 
-The transition between states is triggered by non-parametrized
+The transition between states is triggered by non-parameterized
 actions \texttt{get}, \texttt{move}, and \texttt{put} 
@@
 The fourth component of the type \texttt{igraph}
-assigns security labelled data to locations, a feature not used in the
+assigns security labeled data to locations, a feature not used in the
 current application.
@@
 
-The Privacy protection goal is to avoid deanonymisation. That is, an attacker should not be able to
+The Privacy protection goal is to avoid deanonymization. That is, an attacker should not be able to
 disambiguate the set of pairs of real Ids and EphIds. This is abstractly expressed in the predicate
@@
 ``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
-fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \cite{dp3t:wp}.
+fixed, public string, and PRG is a pseudo-random generator (e.g. AES in counter mode)'' \cite{dp3t:wp}
+%\TODOref{DP-3T Whitepaper}.
 
@@
 to relate any of the \texttt{EphIDi} to $SK_t$  for all \texttt{i} $\in \{1, \ldots, n\}$.
-For a formalisation of this crucial cryptographic property in Isabelle it suffices to define a new type
+For a formalization of this crucial cryptographic property in Isabelle it suffices to define a new type
 of list of Ephemeral Ids \texttt{efidlist} containing the root $SK_t$ (the first \texttt{efid}),
@@
 We define functions for this datatype: \texttt{efidsroot} returning the first of the three
-consituents in an \texttt{efidlist} (the root $SK_t$); \texttt{efids\_index} giving the second
+constituents in an \texttt{efidlist} (the root $SK_t$); \texttt{efids\_index} giving the second
 component, the index of the current \texttt{efid}; \texttt{efids\_inc\_ind} applied to an
@@
 new type \texttt{efidlist}. Note, that in the new datatype \texttt{igraph} this change affects only
-the thrird component, the credentials. The last two componenets, the set of currently present efids,
-and the knowledge set, still operate on the simple type \texttt{efid}.
+the third component, the credentials to become
 \begin{ttbox}
-{\bf datatype} igraph = 
-         Lgraph (location \tttimes location)set 
-                 location \ttfun identity set
    identity \ttfun (string set \tttimes string set \tttimes efidlist)
-                 location \ttfun string \tttimes (dlm \tttimes data) set
-                 location \ttfun efid set
-                 actor \ttfun location \ttfun (identity \tttimes efid) set
 \end{ttbox}  
+The last two components, the set of currently present efids,
+and the knowledge set, remain the same and still operating on the simple type \texttt{efid}.
+%\begin{ttbox}
+%{\bf datatype} igraph = 
+%         Lgraph (location \tttimes location)set 
+%                 location \ttfun identity set
+%
+%                 location \ttfun string \tttimes (dlm \tttimes data) set
+%                 location \ttfun efid set
+%                 actor \ttfun location \ttfun (identity \tttimes efid) set
+%\end{ttbox}
 The refined state transition relation implements the possibility of changing the Ephemeral Ids
-by the rule for the action \texttt{put}.
-\begin{ttbox}
- {\bf{put}}: G = graphI I \ttImp  a \ttatI l \ttImp enables I l (Actor a) put \ttImp
-      I' = Infrastructure (put_graph_efid A l (graphI I))(delta I)
- \ttImp I \ttrelI I' 
-\end{ttbox}
-The change to the infrastructure state is implemented in the function \texttt{put\_graph\_efid} that
+by the rule for the action \texttt{put} that resembles very much the rule for \texttt{get}.
+%\begin{ttbox}
+% {\bf{put}}: G = graphI I \ttImp  a \ttatI l \ttImp enables I l (Actor a) put \ttImp
+%      I' = Infrastructure (put_graph_efid A l (graphI I))(delta I)
+% \ttImp I \ttrelI I' 
+%\end{ttbox}
+The important change to the infrastructure state is implemented in the function \texttt{put\_graph\_efid} that
 increases the current index in the \texttt{efidlist} in the credential component \texttt{cgra g n} for
@@
 by adding an increment (highlighted in the code snippet)
-of the currently used Ephemeral Id before updating the \texttt{egre} component of the target location.
+of the currently used Ephemeral Id before updating the \texttt{egra} component of the target location.
 \begin{ttbox}
@@
 
-\section{Conclusions}
+\section{Summary and discussion of relevance of the approach}
+%{Conclusions}
 \label{sec:concl}
-\subsection{Protection goals of attacks}
-Bluetooth based proximity tracing gives rise to a variety of noteworthy privacy and security issues that deserve formal consideration (and verification).
-The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
-
-In terms of attack scenarios, connecting an Ephemeral Id to a particular device is the first step into multiple different attacks, we want to classify as deanonymisation attacks:
-\begin{itemize}
-\item \textbf{Tracking a device}
-  \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]\\
-  Bluetooth enabled devices introduce some fundamental risks, that are not
-  specific to proximity detection: If the Bluetooth signal is linked to a fixed
-  identifier, say a MAC address, constant broadcasting makes
-  a device trackable.
-  Periodically changing Ephemeral Ids can be an effective mitigation\footnote{%
-    In addition Bluetooth has to be configured to use rotating MAC addresses
-    instead of fixed ones, as is the default in modern smartphone operating
-    systems. Also, rotation of Bluetooth identifiers has to happen
-    in sync with MAC address randomization; the Apple/Google proposal
-    \cite{enf:bluetooth} ensures this.}
-  as they limit the time frame in which a certain id can be used to
-  track the corresponding device, hence, rendering the attack not impossible
-  but significantly less severe.
-\item \textbf{Identify infected individuals}
-  \cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]\\
-  Like any notification about, e.\,g., a STD, might inform a spouse about the
-  infidelity of the other, given the right circumstances, being notified about a
-  possible exposure to COVID-19 might reveal illicit information to
-  the recipient like who it was that tested positive for the virus.
-  This is an inherent vulnerability of any such system,
-  whether it is analogous or digital (digitally aided).\footnote{
-    As DP-3T emphasizes, it it, indeed, inherent to proximity tracing in general
-    \cite{dp3t:psre}[p6].
-    The likelihood of said circumstances to occur, however,
-    might increase when notifications are carried out automatically.
-    Therefore, we feel that it is, nonetheless, worth considering.}
-  In contrast to tracking a Bluetooth enabled device,
-  this attack is not considerably mitigated by changing Ephemeral Ids:
-  %Once connected to a device or individual, an Ephemeral Id might yield the
-  %desired information for up to 14 days
-  A positive test result can be reported for as much as 14 days after the
-  compromised Ephemeral Id was used and still reveal the desired information.
-\end{itemize}
-
-\begin{figure}[htb]
-  \begin{center}
-    \includegraphics[width=6in]{DP-3T_deanonymisation_attacks}
-  \end{center}
-  \caption{Two deanynomisation attacks: Tracking users and identifying confirmed COVID-19 cases}
-  \label{fig:dp3tprot}
-\end{figure}
-
-\subsection{Summary and discussion of relevance of the approach}
+%\subsection{Protection goals of attacks}
+%\subsection
 We can establish in our formal framework an attack that even a system using changing
@@
 attack. In a second refinement step using action refinement based on this insight from the (repeated)
-attack, we can exclude the dangerous attack trace\footnote{Implication on the actual protection goals
-  that are endangered by the attack on privacy are discussed in the larger context of the Corona-warning
-  app (see previous Section)}.
+attack, we can exclude the dangerous attack trace.
+%\footnote{Implication on the actual protection goals
+%  that are endangered by the attack on privacy are discussed in the larger context of the Corona-warning
+%  app (see previous Section)}.
 

### 2020-07-11 paper.tex
one <<<HEAD had to be deleted

@@
 ``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
-fixed, public string, and PRG is a pseudo-random generator (e.g. AES in counter mode)'' \cite{dp3t:wp}
-%\TODOref{DP-3T Whitepaper}.
+fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \cite{dp3t:wp}.
 
@@
 %\subsection
+=======
+\subsection{Protection goals of attacks}
+Bluetooth based proximity tracing gives rise to a variety of noteworthy privacy and security issues that deserve formal consideration (and verification).
+The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, connecting an Ephemeral Id to a particular device is the first step into multiple different attacks, we want to classify as deanonymisation attacks:
+\begin{itemize}
+\item \textbf{Tracking a device}
+  \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]\\
+  Bluetooth enabled devices introduce some fundamental risks, that are not
+  specific to proximity detection: If the Bluetooth signal is linked to a fixed
+  identifier, say a MAC address, constant broadcasting makes
+  a device trackable.
+  Periodically changing Ephemeral Ids can be an effective mitigation\footnote{%
+    In addition Bluetooth has to be configured to use rotating MAC addresses
+    instead of fixed ones, as is the default in modern smartphone operating
+    systems. Also, rotation of Bluetooth identifiers has to happen
+    in sync with MAC address randomization; the Apple/Google proposal
+    \cite{enf:bluetooth} ensures this.}
+  as they limit the time frame in which a certain id can be used to
+  track the corresponding device, hence, rendering the attack not impossible
+  but significantly less severe.
+\item \textbf{Identify infected individuals}
+  \cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]\\
+  Like any notification about, e.\,g., a STD, might inform a spouse about the
+  infidelity of the other, given the right circumstances, being notified about a
+  possible exposure to COVID-19 might reveal illicit information to
+  the recipient like who it was that tested positive for the virus.
+  This is an inherent vulnerability of any such system,
+  whether it is analogous or digital (digitally aided).\footnote{
+    As DP-3T emphasizes, it it, indeed, inherent to proximity tracing in general
+    \cite{dp3t:psre}[p6].
+    The likelihood of said circumstances to occur, however,
+    might increase when notifications are carried out automatically.
+    Therefore, we feel that it is, nonetheless, worth considering.}
+  In contrast to tracking a Bluetooth enabled device,
+  this attack is not considerably mitigated by changing Ephemeral Ids:
+  %Once connected to a device or individual, an Ephemeral Id might yield the
+  %desired information for up to 14 days
+  A positive test result can be reported for as much as 14 days after the
+  compromised Ephemeral Id was used and still reveal the desired information.
+\end{itemize}
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=6in]{DP-3T_deanonymisation_attacks}
+  \end{center}
+  \caption{Two deanynomisation attacks: Tracking users and identifying confirmed COVID-19 cases}
+  \label{fig:dp3tprot}
+\end{figure}
+
+\subsection{Summary and discussion of relevance of the approach}
 We can establish in our formal framework an attack that even a system using changing

### 2020-07-11 submitted_version_paper.tex
submited version

--- 2020-07-11 paper.tex
@@
 %\subsection{Protection goals of attacks}
-%\subsection
+%\subsection{Summary and discussion of relevance of the approach}
 We can establish in our formal framework an attack that even a system using changing

### 2020-07-11 submitted_version_paper.tex
last polish -- is uploaded

@@
 The German Chancellor Angela Merkel has strongly supported the publication of
-the mobile phone Corona warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
+the mobile phone Corona-virus warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
 App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
@@
 
-%The Corona-virus warning app
+%The CWA
 %has taken a long time to develop being published only on
@@
 %
-However, there has been, as of yet, to our knowledge no formal verification been involved.
+However, there has, as of yet, to our knowledge no formal verification been involved.
 Even if a ``post-production'' formal specification seems pointless, 
@@
 been conceived are suitable to cover those weak points,
-or where and to what extend trade offs have to be made due to inherent vulnerabilities.
-Given that the protocol implemented by the framework CWA runs on resembles only the most basic
-of the three proposed by the DP-3T project,
-a formal verified model that stresses the impact or limits of certain security measures
-might help lend some more weight to appeals like \cite{dp3t:impl}
+or where and to what extent trade offs have to be made due to inherent vulnerabilities.
+The protocol implemented by the framework CWA resembles only the most basic
+of the three proposed by the DP-3T project.
+Hence, a formal verified model that stresses the impact or limits of certain security measures
+might give more weight to appeals like \cite{dp3t:impl} 
 %to Apple and Google
@@
 
-The contributions of this paper are (a) formal re-engineering of the Corona-virus warning app
+The contributions of this paper are (a) formal re-engineering of the CWA
 (b) providing an additional security and privacy analysis with interactive theorem proving
@@
 we give a detailed overview of the development history %and security and privacy
-and relevant parts of the Corona-virus warning app (Section \ref{sec:history})
+and relevant parts of the CWA (Section \ref{sec:history})
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
@@
 motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
-framework is introduced and illustrated on the Corona-virus warning app (Section \ref{sec:ref})
+framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
 before drawing some conclusions (Section \ref{sec:concl}).
@@
 core principles of the DP-3T proposal. This API is not only used in the German
-Corona-virus warning app but has the potential of being widely adopted in future
+CWA but has the potential of being widely adopted in future
 app developments that might emerge due to the reach of players like Apple and Google.
@@
   \begin{center}
-  \includegraphics[scale=.35]{theory_structure.pdf}
+  \includegraphics[scale=.45]{theory_structure.pdf}
 \end{center}
@@
 (\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
-for the Corona-virus warning app, express relevant and interesting properties, and conduct
+for the CWA, express relevant and interesting properties, and conduct
 interactive proofs (with the full support of the powerful and highly automated proof support
 of Isabelle).
-%The Corona-virus warning app theory itself is an adaptation of the Infrastructure theory of the 
+%The CWA theory itself is an adaptation of the Infrastructure theory of the 
 %Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
@@
 
-In the Corona-virus warning app, the initial values for the \texttt{igraph} components use
+In the CWA, the initial values for the \texttt{igraph} components use
 two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
@@
 %This will be detailed in the second refinement step
-of the Corona-virus warning app considered next.
+of the CWA considered next.
 

### 2020-07-11 paper.tex
paper: minor fixes and some cleanup

--- 2020-07-11 paper.tex
@@
 Bluetooth Low Energy Beacon protocol.
-The app saves those IDs of people in close proximity.
+The app saves those ids of people in close proximity.
 When at a later date an infected person reports his infection to a central server,
-all clients are provided the means to reconstruct the IDs used by this particular
+all clients are provided the means to reconstruct the ids used by this particular
 person's device over the last 14 days.
-Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
+Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
 %
@@
 In this paper, we first provide some background in Section \ref{sec:background}:
-we give a detailed overview of the development history %and security and privacy
-and relevant parts of the Corona-virus warning app (Section \ref{sec:history})
+we give a brief overview of related works
+and relevant parts of the Corona-virus warning app (Section \ref{sec:history}).
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
@@
 The main reason to focus on this particular family of protocols is the
-\textit{Exposure Notification Framework} (ENF) \cite{enf:proj}, jointly published by Apple and Google, that adopts
+\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google \cite{enf:proj}, that adopts
 %the main
@@
   the backend.}
-\\
 
+\subsubsection{The basic DP-3T architecture}
 All DP-3T protocols roughly work the same:
@@
 %%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
+%\pagebreak
 
@@
 along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
-
-\begin{figure}[htb]
+\begin{figure}
   \begin{center}

### 2020-07-11 paper.tex
reduced 5.1 to an additional subsubsection under 2.1 (to keep the references)

@@
 \end{figure}
+
+\subsubsection{Attack scenarios}
+There is a variety of noteworthy privacy and security issues that deserve formal consideration (and verification). The debate among %emerging between
+advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+As first step, both attacks start with deanonymizing Ephemeral Ids,
+i.\,e. linking them the particular devices.
 
@@
 
-\section{Conclusions}
+\section{Summary and discussion of relevance of the approach}
 \label{sec:concl}
-\subsection{Protection goals of attacks}
-Bluetooth based proximity tracing gives rise to a variety of noteworthy privacy and security issues that deserve formal consideration (and verification).
-The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
-
-In terms of attack scenarios, connecting an Ephemeral Id to a particular device is the first step into multiple different attacks, we want to classify as deanonymisation attacks:
-\begin{itemize}
-\item \textbf{Tracking a device}
-  \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]\\
-  Bluetooth enabled devices introduce some fundamental risks, that are not
-  specific to proximity detection: If the Bluetooth signal is linked to a fixed
-  identifier, say a MAC address, constant broadcasting makes
-  a device trackable.
-  Periodically changing Ephemeral Ids can be an effective mitigation\footnote{%
-    In addition Bluetooth has to be configured to use rotating MAC addresses
-    instead of fixed ones, as is the default in modern smartphone operating
-    systems. Also, rotation of Bluetooth identifiers has to happen
-    in sync with MAC address randomization; the Apple/Google proposal
-    \cite{enf:bluetooth} ensures this.}
-  as they limit the time frame in which a certain id can be used to
-  track the corresponding device, hence, rendering the attack not impossible
-  but significantly less severe.
-\item \textbf{Identify infected individuals}
-  \cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]\\
-  Like any notification about, e.\,g., a STD, might inform a spouse about the
-  infidelity of the other, given the right circumstances, being notified about a
-  possible exposure to COVID-19 might reveal illicit information to
-  the recipient like who it was that tested positive for the virus.
-  This is an inherent vulnerability of any such system,
-  whether it is analogous or digital (digitally aided).\footnote{
-    As DP-3T emphasizes, it it, indeed, inherent to proximity tracing in general
-    \cite{dp3t:psre}[p6].
-    The likelihood of said circumstances to occur, however,
-    might increase when notifications are carried out automatically.
-    Therefore, we feel that it is, nonetheless, worth considering.}
-  In contrast to tracking a Bluetooth enabled device,
-  this attack is not considerably mitigated by changing Ephemeral Ids:
-  %Once connected to a device or individual, an Ephemeral Id might yield the
-  %desired information for up to 14 days
-  A positive test result can be reported for as much as 14 days after the
-  compromised Ephemeral Id was used and still reveal the desired information.
-\end{itemize}
-
-\begin{figure}[htb]
-  \begin{center}
-    \includegraphics[width=6in]{DP-3T_deanonymisation_attacks}
-  \end{center}
-  \caption{Two deanynomisation attacks: Tracking users and identifying confirmed COVID-19 cases}
-  \label{fig:dp3tprot}
-\end{figure}
-
-\subsection{Summary and discussion of relevance of the approach}
 We can establish in our formal framework an attack that even a system using changing

### 2020-07-11 paper.tex
shrunk paper down to 16 pages again :)

--- 2020-07-11 paper.tex
@@
 The app was designed with great
-attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted after a long and
-heated debate with supporters of a central architecture. The distributed architecture is
-based on a very clever application design whereby users' phones broadcast
-highly anonymized so called ``Ephemeral IDs'' at physical locations via the
-Bluetooth Low Energy Beacon protocol.
-The app saves those IDs of people in close proximity.
-When at a later date an infected person reports his infection to a central server,
-all clients are provided the means to reconstruct the IDs used by this particular
-person's device over the last 14 days.
-Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
+attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
+%heated debate with supporters of a central architecture. The distributed architecture is
+that is
+based on a very clever application design whereby clients broadcast
+highly anonymized so called ``Ephemeral IDs'' via Bluetooth and
+store those ids of people in close proximity.
+%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
+%Bluetooth Low Energy Beacon protocol.
+%The app saves those ids of people in close proximity.
+Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
+hence, stored contact data has never to be shared.
+%Infected persons report their infection to a central server,
+%which, in turn, provides all clients the means to reconstruct the ids
+%used by reported cases over the last 14 days.
+%
+%When at a later date an infected person reports his infection to a central server,
+%all clients are provided the means to reconstruct the ids used by this particular
+%person's device over the last 14 days.
+%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
 %
@@
 been conceived are suitable to cover those weak points,
-or where and to what extend trade offs have to be made due to inherent vulnerabilities.
-Given that the protocol implemented by the framework CWA runs on resembles only the most basic
-of the three proposed by the DP-3T project,
+or to what extend trade offs have to be made due to inherent vulnerabilities.
+Given that the protocol implemented by CWA resembles only the most basic
+of the three protocols proposed by the DP-3T project,
 a formal verified model that stresses the impact or limits of certain security measures
-might help lend some more weight to appeals like \cite{dp3t:impl}
+might give more weight to appeals like \cite{dp3t:impl}
 %to Apple and Google
@@
 In this paper, we first provide some background in Section \ref{sec:background}:
-we give a detailed overview of the development history %and security and privacy
-and relevant parts of the Corona-virus warning app (Section \ref{sec:history})
+we give a brief overview of related works
+and the protocol of the Corona-virus warning app (Section \ref{sec:history}).
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
@@
 roof of the \textit{Pan-European Privacy-Preserving Proximity Tracing} project (PEPP-PT) \cite{pepppt:github}, e.g.
-PEPP-PT-ROBERT \cite{pepppt:ROBERT}.
+PEPP-PT-ROBERT \cite{pepppt:ROBERT},
+that might be characterized as centralized architectures.
 
@@
 different protocols with unique properties in terms of privacy and data protection.
-Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
-protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
-DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
-  DP-3T involves a central backend server. It is decentralized with regard to
-  the collection and evaluation of contact information:
-  In centralized architectures the server provides a risk scoring services, whereas decentralized
-  approaches rely on local risk assessment and, thus, do not need to share contact information with
-  the backend.}
-\\
+%Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
+%protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
+%DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
+%  DP-3T involves a central backend server. It is decentralized with regard to
+%  the collection and evaluation of contact information:
+%  In centralized architectures the server provides a risk scoring services, whereas decentralized
+%  approaches rely on local risk assessment and, thus, do not need to share contact information with
+%  the backend.}
 
-All DP-3T protocols roughly work the same:
+\subsubsection{Basic DP-3T architecture}
+%All DP-3T protocols roughly work the same:
 %\subsubsection{Basic DP-3T protocol: How does it work?}
@@
 along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
-
-\begin{figure}[htb]
+\begin{figure}
 \vspace{-.5cm}
   \begin{center}
-    \includegraphics[width=5in]{DP-3T_basic_protocol}
+    \includegraphics[width=4in]{DP-3T_basic_protocol}
   \end{center}
@@
 \end{figure}
+
+\subsubsection{Attack scenarios}
+There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
+The debate among %emerging between
+advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+%As first step, both attacks start with deanonymizing Ephemeral Ids,
+%i.\,e. linking them the particular devices.
+Both attacks rely on deanonymisation of Ephemeral Ids,
+i.\,e. linking ids to particular devices.
 
@@
 %\subsection{Protection goals of attacks}
-%\subsection
-=======
-\subsection{Protection goals of attacks}
-Bluetooth based proximity tracing gives rise to a variety of noteworthy privacy and security issues that deserve formal consideration (and verification).
-The debate emerging between advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
-
-In terms of attack scenarios, connecting an Ephemeral Id to a particular device is the first step into multiple different attacks, we want to classify as deanonymisation attacks:
-\begin{itemize}
-\item \textbf{Tracking a device}
-  \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]\\
-  Bluetooth enabled devices introduce some fundamental risks, that are not
-  specific to proximity detection: If the Bluetooth signal is linked to a fixed
-  identifier, say a MAC address, constant broadcasting makes
-  a device trackable.
-  Periodically changing Ephemeral Ids can be an effective mitigation\footnote{%
-    In addition Bluetooth has to be configured to use rotating MAC addresses
-    instead of fixed ones, as is the default in modern smartphone operating
-    systems. Also, rotation of Bluetooth identifiers has to happen
-    in sync with MAC address randomization; the Apple/Google proposal
-    \cite{enf:bluetooth} ensures this.}
-  as they limit the time frame in which a certain id can be used to
-  track the corresponding device, hence, rendering the attack not impossible
-  but significantly less severe.
-\item \textbf{Identify infected individuals}
-  \cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]\\
-  Like any notification about, e.\,g., a STD, might inform a spouse about the
-  infidelity of the other, given the right circumstances, being notified about a
-  possible exposure to COVID-19 might reveal illicit information to
-  the recipient like who it was that tested positive for the virus.
-  This is an inherent vulnerability of any such system,
-  whether it is analogous or digital (digitally aided).\footnote{
-    As DP-3T emphasizes, it it, indeed, inherent to proximity tracing in general
-    \cite{dp3t:psre}[p6].
-    The likelihood of said circumstances to occur, however,
-    might increase when notifications are carried out automatically.
-    Therefore, we feel that it is, nonetheless, worth considering.}
-  In contrast to tracking a Bluetooth enabled device,
-  this attack is not considerably mitigated by changing Ephemeral Ids:
-  %Once connected to a device or individual, an Ephemeral Id might yield the
-  %desired information for up to 14 days
-  A positive test result can be reported for as much as 14 days after the
-  compromised Ephemeral Id was used and still reveal the desired information.
-\end{itemize}
-
-\begin{figure}[htb]
-  \begin{center}
-    \includegraphics[width=6in]{DP-3T_deanonymisation_attacks}
-  \end{center}
-  \caption{Two deanynomisation attacks: Tracking users and identifying confirmed COVID-19 cases}
-  \label{fig:dp3tprot}
-\end{figure}
-
-\subsection{Summary and discussion of relevance of the approach}
 We can establish in our formal framework an attack that even a system using changing

### 2020-07-11 paper.tex
moved attacks subsection to improve vertical distribution - fig. are not perfectly placed, though :(
merged paper.tex with submitted_version.tex (missing: replacement of Corona-virus warning app with CWA)

@@
 been conceived are suitable to cover those weak points,
-or to what extend trade offs have to be made due to inherent vulnerabilities.
-Given that the protocol implemented by CWA resembles only the most basic
-of the three protocols proposed by the DP-3T project,
-a formal verified model that stresses the impact or limits of certain security measures
+or to what extent trade offs have to be made due to inherent vulnerabilities.
+The protocol implemented by the framework CWA resembles only the most basic
+of the three protocols proposed by the DP-3T project.
+Hence, a formal verified model that stresses the impact or limits of certain security measures
 might give more weight to appeals like \cite{dp3t:impl}
@@
 we give a brief overview of related works
-and the protocol of the Corona-virus warning app (Section \ref{sec:history}).
+and the protocol of the German app (Section \ref{sec:history}).
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
@@
 
+There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
+The debate among %emerging between
+advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+%As first step, both attacks start with deanonymizing Ephemeral Ids,
+%i.\,e. linking them the particular devices.
+%Both attacks rely on deanonymisation of Ephemeral Ids,
+%i.\,e. linking ids to particular devices.
+
 \subsubsection{Basic DP-3T architecture}
@@
 along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
-\begin{figure}
+\begin{figure}[htb]
 \vspace{-.5cm}
@@
 
-\subsubsection{Attack scenarios}
-There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
-The debate among %emerging between
-advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+%\subsubsection{Attack scenarios}
+%There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
+%The debate among %emerging between
+%advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
 
-In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
-%As first step, both attacks start with deanonymizing Ephemeral Ids,
-%i.\,e. linking them the particular devices.
-Both attacks rely on deanonymisation of Ephemeral Ids,
-i.\,e. linking ids to particular devices.
+%In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+%%As first step, both attacks start with deanonymizing Ephemeral Ids,
+%%i.\,e. linking them the particular devices.
+%Both attacks rely on deanonymisation of Ephemeral Ids,
+%i.\,e. linking ids to particular devices.
 
@@
 %\subsection{Protection goals of attacks}
+%\subsection{Summary and discussion of relevance of the approach}
 We can establish in our formal framework an attack that even a system using changing

### 2020-07-11 paper.tex
increase image size (basic protocol) as it does no harm positioning-wise
change my email address
cleanup: removed old (commented out) text snippets - only in "my" parts

@@
 %\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
-%\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
-%\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
 \newcommand\ttand{\mbox{{$\land$}}}
@@
 \institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
-\email{f.kammueller@mdx.ac.uk|bialut@gmail.com}
+\email{f.kammueller@mdx.ac.uk|bianca.lutz@campus.tu-berlin.de}
 }
@@
 conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
-%The Corona-virus app has been produced based on a sophisticated security
-%concept conceived by experts in the field %to our knowledge, no formal verification
-%%
-%and has strong claims with regard to mathematical support:
-%
-%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
-%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
 %
@@
 might give more weight to appeals like \cite{dp3t:impl}
-%to Apple and Google
 to adopt the more sophisticated protocols.
-%
-%Thereby, we believe
-%that our current work is useful to increase the trust in the Corona-virus app necessary for
-%its wide adoption which in turn is crucial for it to be efficient.
 
@@
 There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
-The debate among %emerging between
-advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+The debate among advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
 
 In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
-%As first step, both attacks start with deanonymizing Ephemeral Ids,
-%i.\,e. linking them the particular devices.
-%Both attacks rely on deanonymisation of Ephemeral Ids,
-%i.\,e. linking ids to particular devices.
 
 \subsubsection{Basic DP-3T architecture}
-%All DP-3T protocols roughly work the same:
-%\subsubsection{Basic DP-3T protocol: How does it work?}
-%  How does it relate to ENF/CWA? (broad strokes)}
-%
 Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
@@
 privacy and data protection properties to different degrees but are more costly to implement.
-%Aside from additional security measures,
-%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
-%like signing content the server provides as a general rule, the protocol implemented by ENF and,
-%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
-%privacy and data protection properties to different degrees but are more costly.
-%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
-%\pagebreak
-
+%
 Figure \ref{fig:dp3tprot} illustrates the basic system architecture
@@
   \begin{center}
-    \includegraphics[width=4in]{DP-3T_basic_protocol}
+    \includegraphics[width=5in]{DP-3T_basic_protocol}
   \end{center}
@@
 \end{figure}
-
-%\subsubsection{Attack scenarios}
-%There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
-%The debate among %emerging between
-%advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
-
-%In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
-%%As first step, both attacks start with deanonymizing Ephemeral Ids,
-%%i.\,e. linking them the particular devices.
-%Both attacks rely on deanonymisation of Ephemeral Ids,
-%i.\,e. linking ids to particular devices.
 

### 2020-07-11 submitted_version_paper.tex
new (submitted) version: fixed citation

@@
 The app was designed with great
-attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted after a long and
-heated debate with supporters of a central architecture. The distributed architecture is
-based on a very clever application design whereby users' phones broadcast
-highly anonymized so called ``Ephemeral IDs'' at physical locations via the
-Bluetooth Low Energy Beacon protocol.
-The app saves those IDs of people in close proximity.
-When at a later date an infected person reports his infection to a central server,
-all clients are provided the means to reconstruct the IDs used by this particular
-person's device over the last 14 days.
-Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
+attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
+%heated debate with supporters of a central architecture. The distributed architecture is
+that is
+based on a very clever application design whereby clients broadcast
+highly anonymized so called ``Ephemeral Ids'' via Bluetooth and
+store those ids of people in close proximity.
+%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
+%Bluetooth Low Energy Beacon protocol.
+%The app saves those ids of people in close proximity.
+Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
+hence, stored contact data has never to be shared.
+%Infected persons report their infection to a central server,
+%which, in turn, provides all clients the means to reconstruct the ids
+%used by reported cases over the last 14 days.
+%
+%When at a later date an infected person reports his infection to a central server,
+%all clients are provided the means to reconstruct the ids used by this particular
+%person's device over the last 14 days.
+%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
 %
@@
 been conceived are suitable to cover those weak points,
-or where and to what extent trade offs have to be made due to inherent vulnerabilities.
+or to what extent trade offs have to be made due to inherent vulnerabilities.
 The protocol implemented by the framework CWA resembles only the most basic
-of the three proposed by the DP-3T project.
+of the three protocols proposed by the DP-3T project.
 Hence, a formal verified model that stresses the impact or limits of certain security measures
 might give more weight to appeals like \cite{dp3t:impl}
-%to Apple and Google
 to adopt the more sophisticated protocols.
-%
-%Thereby, we believe
-%that our current work is useful to increase the trust in the Corona-virus app necessary for
-%its wide adoption which in turn is crucial for it to be efficient.
 
@@
 In this paper, we first provide some background in Section \ref{sec:background}:
-we give a detailed overview of the development history %and security and privacy
-and relevant parts of the CWA (Section \ref{sec:history})
+we give a brief overview of related works
+and the protocol of the German app (Section \ref{sec:history}).
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
@@
 different protocols with unique properties in terms of privacy and data protection.
-Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
-protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
-DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
-  DP-3T involves a central backend server. It is decentralized with regard to
-  the collection and evaluation of contact information:
-  In centralized architectures the server provides a risk scoring services, whereas decentralized
-  approaches rely on local risk assessment and, thus, do not need to share contact information with
-  the backend.}
-\\
+%Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
+%protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
+%DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
+%  DP-3T involves a central backend server. It is decentralized with regard to
+%  the collection and evaluation of contact information:
+%  In centralized architectures the server provides a risk scoring services, whereas decentralized
+%  approaches rely on local risk assessment and, thus, do not need to share contact information with
+%  the backend.}
 
-All DP-3T protocols roughly work the same:
-%\subsubsection{Basic DP-3T protocol: How does it work?}
-%  How does it relate to ENF/CWA? (broad strokes)}
-%
+There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
+The debate among advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.
+
+In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+
+\subsubsection{Basic DP-3T architecture}
 Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
@@
 privacy and data protection properties to different degrees but are more costly to implement.
-%Aside from additional security measures,
-%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
-%like signing content the server provides as a general rule, the protocol implemented by ENF and,
-%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
-%privacy and data protection properties to different degrees but are more costly.
-%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
-
+%
 Figure \ref{fig:dp3tprot} illustrates the basic system architecture
 along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
-
 \begin{figure}[htb]

### 2020-07-11 paper.tex
"the German app" -> "CWA"

@@
 based on a very clever application design whereby clients broadcast
-highly anonymized so called ``Ephemeral IDs'' via Bluetooth and
+highly anonymized so called ``Ephemeral Ids'' via Bluetooth and
 store those ids of people in close proximity.
@@
 we give a brief overview of related works
-and the protocol of the German app (Section \ref{sec:history}).
+and the protocol of CWA (Section \ref{sec:history}).
 We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).

### 2020-07-11 submitted_version_paper.tex
copied from paper

@@
 We provide a model in the Isabelle Infrastructure framework of the recently published German
-Corona-virus warning app. The app supports breaking infection chains by informing users
+Corona-virus warning app (CWA). The app supports breaking infection chains by informing users
 whether they have been in close contact to an infected person. The app has a decentralized
@@
 refinement process of the Isabelle Infrastructure framework to highlight how the use of
-continuously changing ephemeral ids improves the anonymity.
+continuously changing Ephemeral Ids (EphIDs) improves the anonymity.
 \end{abstract}
@@
 based on a very clever application design whereby clients broadcast
-highly anonymized so called ``Ephemeral Ids'' via Bluetooth and
-store those ids of people in close proximity.
+highly anonymized identifiers (ids)
+%so called ``Ephemeral Ids''
+via Bluetooth and store those ids of people in close proximity.
 %highly anonymized so called ``Ephemeral IDs'' at physical locations via the
@@
 based on attacks found by the attack tree analysis and (d) an application of the refinement to
-improve security of the Corona-virus app specification.
+improve security of the CWA specification.
 
@@
 \subsubsection{Basic DP-3T architecture}
-Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
+Upon installation, the app generates secret daily seeds to derive so-called \textit{Ephemeral Ids}
 (EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
@@
 \vspace{-.5cm}
-\caption{Generic Isabelle Infrastructure framework applied to Corona-warning app.}
+\caption{Generic Isabelle Infrastructure framework applied to CWA.}
 \label{fig:theorystruc}
@@
 \subsubsection{Instantiation of Framework}
-The formal model of the Corona-warning app uses the Isabelle Infrastructure framework instantiating it
+The formal model of CWA uses the Isabelle Infrastructure framework instantiating it 
 by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
@@
 
-\section{Modeling and analyzing Corona-warning app}
+\section{Modeling and analyzing CWA}
 \label{sec:model}
@@
 \end{ttbox}
-In the current application of the framework to the Corona-warning app case study, 
+In the current application of the framework to the CWA case study, 
 this graph contains a set of location pairs representing the %structure 
@@
 most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
-just a natural number but will be refined to actually represent lists of ``Ephemeral'' Ids later
+just a natural number but will be refined to actually represent lists of Ephemeral Ids later
 when refining the specification.
@@
 specified in the set which is the second element of that pair.
-The local policies definition for the Corona-warning app, simply permits all actions to
+The local policies definition for the CWA, simply permits all actions to
 all actors in both locations.
@@
 \end{ttbox}  
-For the Corona-warning app, the initial infrastructure contains the graph \texttt{ex\_graph}
+For the CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
 with its two locations pub and shop and is then wrapped up with the local policies
@@
 The Privacy protection goal is to avoid deanonymization. That is, an attacker should not be able to
-disambiguate the set of pairs of real Ids and EphIds. This is abstractly expressed in the predicate
+disambiguate the set of pairs of real ids and EphIDs. This is abstractly expressed in the predicate
 identifiable.
@@
 \begin{ttbox}
-  EphId1 || ... || EphIdn = PRG(PRF(SKt,``broadcast key''))
+  EphID1 || ... || EphIDn = PRG(PRF(SKt,``broadcast key''))
 \end{ttbox} 
@@
 identified on the very abstract level (fixed ids) persists.
-The refinement allows refining the datatype (Id $\mapsto$ EphId) but also delivers the usual trace
+The refinement allows refining the datatype (Id $\mapsto$ EphID) but also delivers the usual trace
 refinement (behaviours of the refined systems are a subset of the traces of the abstract system).

### 2020-07-11 paper.tex
update

@@
 The German Chancellor Angela Merkel has strongly supported the publication of
-the mobile phone Corona-virus warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
-App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
+the mobile phone Corona-virus warning app (CWA; \cite{cwa:github}) by publicly proclaiming that
+``this App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
 in Germany have downloaded the app with 6 million on the first day.

### 2020-07-11 paper.tex
the CWA -> CWA

@@
 
-The contributions of this paper are (a) formal re-engineering of the CWA
+The contributions of this paper are (a) formal re-engineering of CWA
 (b) providing an additional security and privacy analysis with interactive theorem proving
@@
 motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
-framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
+framework is introduced and illustrated on CWA (Section \ref{sec:ref})
 before drawing some conclusions (Section \ref{sec:concl}).
@@
 %the main
-core principles of the DP-3T proposal. This API is not only used in the German
+core principles of the DP-3T proposal. This API is not only used in 
 CWA but has the potential of being widely adopted in future
@@
 (\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
-for the CWA, express relevant and interesting properties, and conduct
+for CWA, express relevant and interesting properties, and conduct
 interactive proofs (with the full support of the powerful and highly automated proof support
@@
 
-In the CWA, the initial values for the \texttt{igraph} components use
+In CWA, the initial values for the \texttt{igraph} components use
 two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
@@
 specified in the set which is the second element of that pair.
-The local policies definition for the CWA, simply permits all actions to
+The local policies definition for CWA, simply permits all actions to
 all actors in both locations.
@@
 \end{ttbox}  
-For the CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
+For CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
 with its two locations pub and shop and is then wrapped up with the local policies
@@
 %This will be detailed in the second refinement step
-of the CWA considered next.
+of CWA considered next.
 

### 2022-02-28 DPM2020/final/paper.tex
moved the folder final into DPM2020 as its relevant there

--- 2020-07-11 paper.tex
@@
 Even if a ``post-production'' formal specification seems pointless, 
-it allows to reveal weak points of the architecture, show that the measures that have
+it allows revealing weak points of the architecture, show that the measures that have
 been conceived are suitable to cover those weak points,
-or to what extent trade offs have to be made due to inherent vulnerabilities.
+or to what extent trade-offs have to be made due to inherent vulnerabilities.
 The protocol implemented by the framework CWA resembles only the most basic
@@
 certification of a novel view on the system architecture including actors, locations, and policies,
-(c) a formal definition of a security refinement process that allows to improve a system
+(c) a formal definition of a security refinement process that allows improving a system
 based on attacks found by the attack tree analysis and (d) an application of the refinement to
@@
 The formal model in the Isabelle insider framework is fully mechanized and proved in
-Isabelle (sources available \cite{kam:18smc}).
+Isabelle (sources available \cite{kam:20gitat}).
 
@@
 
-In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
+In terms of attack scenarios, we focus on, what might be classified as deanonymization attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
 
@@
 \item {Kripke structures and state transitions:}\\ 
-Using a generic state transition relation $\mapsto$, Kripke structures
-are defined as a set of states reachable by $\mapsto$ from an initial
-state set, for example
+Using a generic state transition relation $\to$, Kripke structures
+are defined as a set of states reachable by $\to$ from an initial
+state set, for example,
 \begin{ttbox}
@@
 \end{ttbox}
+where $\to^*$ is the reflexive transitive closure over $\to$.
 \item {CTL statements:}\\ 
@@
 system refinement presented in this paper as introduced in the next section).
-An abstract attack tree may be refined by spelling out the attack steps until a valid attack
-is reached:
+An abstract attack tree \texttt{A} of type \texttt{\ttalpha\ attree} over an arbitrary state type \texttt{\ttsigma}
+may be refined by spelling out the attack steps until a valid attack is reached:
 
-\texttt{\ttvdash A :: (\ttsigma :: state) attree)}.
+\texttt{\ttvdash A :: (\ttsigma :: state) attree}.
 
@@
 and can be used to facilitate actual application verification as demonstrated 
-her in the stepwise system refinements.
+here in the stepwise system refinements.
 \end{itemize}
@@
 by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
-to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
+to physical locations. The Ephemeral Ids, their sending and change is added to infrastructures
 by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
-for the actions move, get, and put. The details of the newly adapted Infrastructure are
+for the actions move, get, and put. The details of the newly adapted infrastructure are
 presented in Section \ref{sec:model}.
@@
 %In the remainder of this paper, we introduce the model that we conceived for . 
-All Isabelle sources are available online \cite{kam:20gitsc}.
+All Isabelle sources are available online \cite{kam:20gitat}.
 
@@
 formal definition of a combination of trace refinement and structural 
-refinement (or datatype refinement). The definition allows to prove property preservation results 
+refinement (or datatype refinement). The definition allows proving property preservation results 
 crucial for an iterative development process.
@@
 topology of the infrastructure
-as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
+as a graph of nodes and a function %\footnote{ We use the common $\lambda$-abstraction, 
 %e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
@@
 most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
-just a natural number but will be refined to actually represent lists of Ephemeral Ids later
+just a natural number of type \texttt{nat} mapped by the constructor \texttt{Efid} to the new datatype
+\texttt{efid} but will be refined to actually represent lists of Ephemeral Ids later
 when refining the specification.
@@
 disambiguate the set of pairs of real ids and EphIDs. This is abstractly expressed in the predicate
-identifiable.
+\texttt{identifiable}.
 \begin{ttbox}
@@
 \end{ttbox}
-The predicate identifiable is used to express as the global policy `Eve cannot deanonymize an Ephemeral
+The predicate \texttt{identifiable} is used to express as the global policy `Eve cannot deanonymize an Ephemeral
 Id \texttt{eid} using the gathered knowledge':
@@
 \label{sec:statetrans}
-The state transition relation uses the syntactic infix notation 
-\texttt{I \ttrelI\, I'}  to denote that infrastructures 
-\texttt{I} and \texttt{I'} are in this relation.
-%Rules for the put, get, move, process and delete actions exist.
-To give an impression of this definition, we show first the rule defining the state transition for
-the action get. Initially, this rule expresses that
-an actor that resides at a location \texttt{l} (\texttt{a \ttatI\ l})
-and is enabled by the local policy in this location to ``get'' can
-combine  all ids at the current location (contained in \texttt{egra G l}) with
-all actors at the current location (contained in \texttt{agra G l}) and add this
-set of pairs to his knowledge set \texttt{kgra G} using the function update \texttt{f(l := n)}
-redefining the function \texttt{f} for the input \texttt{l} to have now the new value \texttt{n}.
-\begin{ttbox}
- {\bf{get\_data}}: G = graphI I \ttImp a \ttatI l \ttImp enables I l (Actor a) get \ttImp 
-   I' = Infrastructure 
-          (Lgraph (gra G)(agra G)(cgra G)(lgra G)(egra G)
-            ((kgra G)((Actor a) := ((kgra G (Actor a))(l:=
-                 \{(x,y). x \ttin agra G l \ttand y \ttin egra G l\})))))
-          (delta I)
-   \ttImp I \ttrel{n} I' 
-\end{ttbox}
-Another interesting rule for the state transition is the one for \texttt{move} whose structure
-resembles the previous one.
+The state transition relation uses the syntactic infix notation \texttt{I \ttrel\, I'}  to denote that
+infrastructures \texttt{I} and \texttt{I'} are in this relation. 
+We will use this state transition to define the dynamic behaviour of the system including actors' movements and
+the evolution of their knowledge. This allows representing attacks as sequences of state transitions leading to
+states in which the global policy is violated. 
+To give an intuitive motivating idea before the formal presentation in Section \ref{sec:ana}: the attack will
+consist of the state transitions of Eve getting available EphIds at the location pub, Alice moving to shop, Eve
+following her and again getting the EphIds at shop.
+
+So, one relevant rule for the state transition is the one for \texttt{move} that defines the state transition relation
+expressing state changes when actors move across the infrastructure.
+%whose structure resembles the previous one.
 \begin{ttbox}
@@
     I' = Infrastructure (move_graph_a a l l' (graphI I))(delta I)
- \ttImp I \ttrelI I' 
+ \ttImp I \ttrel I' 
 \end{ttbox}
@@
          (kgra g)
+\end{ttbox}  
+%Rules for the put, get, move, process and delete actions exist.
+%To give an impression of this definition,
+Another action relevant for the deanonymization attack describes how an actor gets information.
+This is defined by the rule of the state transition for the action get. Initially, this rule
+expresses that an actor that resides at a location \texttt{l} (\texttt{a \ttatI\ l})
+and is enabled by the local policy in this location to ``get'' can
+combine  all ids at the current location (contained in \texttt{egra G l}) with
+all actors at the current location (contained in \texttt{agra G l}) and add this
+set of pairs to his knowledge set \texttt{kgra G} using the function update \texttt{f(l := n)}
+redefining the function \texttt{f} for the input \texttt{l} to have now the new value \texttt{n}.
+\begin{ttbox}
+ {\bf{get\_data}}: G = graphI I \ttImp a \ttatI l \ttImp enables I l (Actor a) get \ttImp 
+   I' = Infrastructure 
+          (Lgraph (gra G)(agra G)(cgra G)(lgra G)(egra G)
+            ((kgra G)((Actor a) := ((kgra G (Actor a))(l:=
+                 \{(x,y). x \ttin agra G l \ttand y \ttin egra G l\})))))
+          (delta I)
+   \ttImp I \ttrel I' 
 \end{ttbox}
@@
 \end{ttbox}
-We can simply apply the Correctness theorem \texttt{AT\_EF} to 
-immediately prove this CTL statement. This application of the meta-theorem 
+We can simply apply the Correctness theorem \texttt{AT\_EF} \cite{kam:18b,kam:18a}
+\begin{ttbox}
+\ttvdash A \ttImp attack A = (I,s)
+\ttImp Kripke \{s . \ttexists i \ttin I. (i \ttrelIstar s) \} I \ttvdash EF s
+\end{ttbox}  
+to immediately prove this CTL statement. This application of the meta-theorem 
 of Correctness of attack trees saves us proving the CTL formula tediously 
@@
 \label{sec:ref}
-%\TODO{This section is just copied over from arxive paper - adapt}
 Intuitively, a refinement changes some aspect of the type of
@@
 $\sqsubseteq_{\mathcal{E}}$. 
-The relation \texttt{mod\_trans} is typed as a relation over triples --
+This relation, called \texttt{refinement}, is typed as a relation over triples --
 a function from a threefold Cartesian product to \texttt{bool}, the 
@@
 Consequently, the middle element of the triples selected by the relation 
-\texttt{mod\_trans} is a function of type $\sigma' \Rightarrow \sigma$ 
+\texttt{refinement} is a function of type $\sigma' \Rightarrow \sigma$ 
 mapping elements of the refined state to the abstract state.
@@
 infix syntax in Isabelle that allows the definition of mathematical notation
-instead of writing \texttt{mod\_trans} in prefix manner. This nicer infix
+instead of writing \texttt{refinement} in prefix manner. This nicer infix
 syntax is already used in the actual definition.
@@
 They are logically equivalent but of different
-types: within a HOL formula $P$, for example, as below $\forall x. P \ttimp Q$, only the implication $\ttimp$
-can be used.
+types: within a HOL formula, for example, within the below $\forall $\texttt{s' \ttin\ states K'.}\dots,
+only the implication $\ttimp$ can be used.
 \begin{ttbox}
@@
 $s$ must also be an initial state and from there the image of $s'$
-under ${\mathcal E}$ must be reached with $0$ or $n$ steps.
+under ${\mathcal E}$ must be reached within $0$ and $n$ steps.
 
@@
 where the operator \texttt{$\ttimg$} in \texttt{\ttecal\ttimg(init K')}
-represents function image, that is the set, $\{\ttecal(x). x \in \texttt{init K'}\} $.
+represents function image, i.e.,  $\{\ttecal(x) \mid x \in \texttt{init K'}\} $.
 \begin{ttbox}
@@
 It is remarkable, that our definition of refinement by Kripke 
-structure refinement entails property preservation and makes it possible 
+structures entails property preservation and makes it possible 
 to prove this as a theorem in Isabelle once for all, i.e., as a meta-theorem.
 However, this is due to the fact that our generic definition of state transition
-allows to explicitly formalize such sophisticated concepts like reachability.
+allows explicitly formalizing such sophisticated concepts like reachability.
 For practical purposes, however, the proof obligation of showing that
@@
 but neither $s$ nor $s'$ are reachable from initial states in \texttt{K'}.
-For cases, where we want to have the simpler one-step proviso but still need 
-reachability we provide a slightly weaker version of \texttt{strong\_mt}.
+For cases, where we %want to
+have the simpler one-step proviso %but still need 
+and also reachability from initial states 
+we provide a slightly weaker version of \texttt{strong\_mt}.
+%that pre-supposes reachability of \texttt{s} from initial states.
 \begin{ttbox}
@@
 trace refinement as it is given in process algebras like CSP. In this view,
-the properties of a system are given by the set of its traces. Now, a refinement
-of the system is given by another system that has a subset of the traces of the 
+the properties of a system are given by the set of its event traces. Now, a refinement
+of the system is given by another system that has a subset of the event traces of the 
 former one.
 Although the principal idea is similar, we greatly extend it since our notion
-additionally incorporates refinement. Since we include a state map 
+additionally incorporates data refinement. Since we include a state map 
 \texttt{\ttsigma'\ttfun \ttsigma} in our refinement map, we additionally
@@
 From a cryptographic point of view, the crucial properties of the Ephemeral Ids are that
-they are purely random, therefore, they cannot be guessed, but at the same time if -- after
+they are purely random, therefore, they cannot be guessed, but, at the same time, if -- after
 the actual encounter between sender and receiver -- the seed $SK_t$ is published, it is feasible
@@
 We define functions for this datatype: \texttt{efidsroot} returning the first of the three
-constituents in an \texttt{efidlist} (the root $SK_t$); \texttt{efids\_index} giving the second
-component, the index of the current \texttt{efid}; \texttt{efids\_inc\_ind} applied to an
+constituents in an \texttt{efidlist} (the root $SK_t$ of type \texttt{efid}); \texttt{efids\_index} giving the second
+component of type \texttt{nat}, that is, the index of the current \texttt{efid}; \texttt{efids\_inc\_ind} applied to an
 \texttt{efidlist} increments the index; \texttt{efids\_cur} returning the current \texttt{efid}
-from the list and \texttt{efids\_list} for the entire list (the third component).
+from the list and \texttt{efids\_list} for the entire list (the third component of type \texttt{efid list}).
 
@@
 The last two components, the set of currently present efids,
-and the knowledge set, remain the same and still operating on the simple type \texttt{efid}.
+and the knowledge set, remain the same and still operate on the simple type \texttt{efid}.
 %\begin{ttbox}
@@
 
-Surprisingly, we can still prove \texttt{corona\_KripkeO \ttvdash EF scoronaO} by using the
-same attack tree as in the abstract model: if Bob moves from pub to shop, he is vulnerable to
+Surprisingly, we can still prove \texttt{corona\_KripkeO \ttvdash\ EF scoronaO} by using the
+same attack tree as in the abstract model: if Bob moves from pub to shop\footnote{that is, if he moves alone:
+  if all others from pub go to shop with him his anonymity remains intact.}, he is vulnerable to
 being identifiable as long as he does not change the current \texttt{efid}. So, if Eve moves to
-the shop as well and performs a get before Bob does a put, then Eve's knowledge sets permits identifying
+the shop as well and performs a get before Bob does a put, then Eve's knowledge sets permit identifying
 Bob's current Ephemeral Id as his.
@@
 We can implement that change by a minimal update to the function \texttt{move\_graph\_a} (see Section \ref{sec:model})
-by adding an increment (highlighted in the code snippet)
+by adding an increment (highlighted in the code snippet in bold letters)
 of the currently used Ephemeral Id before updating the \texttt{egra} component of the target location.
@@
 omitting a few steps (the refinement relation is defined using the reflexive transitive closure
-\texttt{\ttrelIstar}.
+\texttt{\ttrelIstar}).
 
@@
 When establishing the attack we start from a simplistic scenario that does not use Ephemeral Ids but fixed ids.
-In this (over)simplified model the attack is established. We then define a formal Refinement calculus for
+In this (over)simplified model the attack is established. We then define a formal refinement calculus for
 the Isabelle Infrastructure framework to refine the attack to a system with Ephemeral Ids that change in fixed
@@
 step binds the action move together with a put action. %which effectively represents an action refinement.
-This shows that besides datatype refinement and trace refinement our Refinement calculus also entails action
+This shows that besides datatype refinement and trace refinement our refinement calculus also entails action
 refinement. This action refinement is implemented implicitly by changing the effects of the actions
